### 2019-07-17 metric-learn-jmlr.tex
wip

\documentclass[twoside,11pt]{article}

% Any additional packages needed should be included after jmlr2e.
% Note that jmlr2e.sty includes epsfig, amssymb, natbib and graphicx,
% and defines many common macros, such as 'proof' and 'example'.
%
% It also sets the bibliographystyle to plainnat; for more information on
% natbib citation styles, see the natbib documentation, a copy of which
% is archived at http://www.jmlr.org/format/natbib.pdf

\usepackage{jmlr2e}
\usepackage{minted}
\usepackage[colorinlistoftodos]{todonotes}

% Definitions of handy macros can go here

\newcommand{\dataset}{{\cal D}}
\newcommand{\fracpartial}[2]{\frac{\partial #1}{\partial  #2}}

% Heading arguments are {volume}{year}{pages}{submitted}{published}{author-full-names}

\jmlrheading{1}{2000}{1-48}{4/00}{10/00}{Marina Meil\u{a} and Michael I. Jordan}

% Short headings should be running head and authors last names

\ShortHeadings{Learning with Mixtures of Trees}{Meil\u{a} and Jordan}
\firstpageno{1}

\begin{document}

\title{Metric-learn: Metric Learning in Python}

\author{\name Marina Meil\u{a} \email mmp@stat.washington.edu \\
       \addr Department of Statistics\\
       University of Washington\\
       Seattle, WA 98195-4322, USA
       \AND
       \name Michael I.\ Jordan \email jordan@cs.berkeley.edu \\
       \addr Division of Computer Science and Department of Statistics\\
       University of California\\
       Berkeley, CA 94720-1776, USA}

\editor{Leslie Pack Kaelbling}

\maketitle

\begin{abstract}%   <- trailing '%' for backward compatibility of .sty file
TODO
\end{abstract}

\begin{keywords}
  Machine Learning, Python, Metric Learning, Scikit-learn
\end{keywords}

\section{Introduction}

\begin{itemize}
    \item Metric learning
    \item Importance of supporting the classic ML workflows for model selection, model evaluation, pipelining with other estimators etc, via sklearn. makes metric-learn stands out from other packages
    in particular in the case of weakly supervised algorithms
\end{itemize}

% Acknowledgements should go at the end, before appendices and references

\section{Overview of the Package}
\texttt{metric-learn} is  a
\todo[inline]{GitHub repo, PyPi, mention current release, license}
\texttt{metric-learn} only relies on core libraries from the SciPy ecosystem: \texttt{numpy}, \texttt{scipy}, and \texttt{scikit-learn}. An optional dependency with \texttt{skggm} (\cite{skggm}) can be installed in order to use a more stable solver when using the LSML algorithm from the package.
It can be installed from the Python Package Index (PyPI), on python 2.7 and 3.5 or later.  The source code is available at \url{http://github.com/metric-learn/metric-learn}. The development is collaborative and open to all. The code is free of use, provided under the MIT licence. The current realease is 0.5.0. It includes many refactoring including the ones described in the paper, that improve scikit-learn compatibility.

\section{Available algorithms}

\texttt{metric-learn} contains most of the highly cited Mahalanobis metric learning algorithms. Here is a list of the algorithms contained in \texttt{metric-learn}:
\begin{itemize}
    \item Supervised Metric Learning algorithms: 
    Supervised algorithms learns a metric that brings closer together points from the same class and spreads further away points from a different class.
    \begin{itemize}
        \item NCA: Neighborhood Components Analysis
        \item LMNN: Large Margin Nearest Neighbors
        \item RCA: Relative Components Analysis
        \item LFDA: Local Fisher Discriminant Analysis
    \end{itemize}
    \item Weakly Supervised Metric Learning algorithms:
    Weakly Supervised algorithms use less information than the previous algorihtms: they only take tuple of points and possibly a corresponding label. See below for more concrete examples of algorithms learning on tuples.
    \begin{itemize}
        \item Pairs Learners: 
        Pairs learners take as input pair of points, and a corresponding label indicating whether the two points are similar or not. They will learn a metric that brings points from a similar pair closer together, and points from a dissimilar pair further away from each other.
        \begin{itemize}
            \item MMC: Mahalanobis Metric for Clustering
            \item SDML: Sparse High-Dimensional Metric Learning (SDML)
            \item ITML: Information Theoretic Metric Learning (ITML)
        \end{itemize}
        \item Quadruplets Learners: 
        Quadruplets learners take as input quadruplets of points. They learn a metric that brings the two first points of each quadruplet closer than the two last points are.
        \begin{itemize}
            \item LSML: Metric Learning from Relative Comparisons by Minimizing Squared Residual
        \end{itemize}
    \end{itemize}
\end{itemize}

\todo[inline]{Mention and link documentation}

\textbf{Documentation} A detailed documentation, including installation guidelines, the description of the algorithms and the API, as well as example of use done with \texttt{sphinx-gallery}, is available at \url{https://metric-learn.github.io/metric-learn/}. It also includes \texttt{intersphinx} bindings for more efficient navigation through the documentation of the whole \texttt{scipy} ecoystem.

\todo[inline]{Code example to showcase usage, potentially with compelling visualization?}

\todo[inline]{what about four plots: two plots for supervised metric learning: before and after learning the metric (classes are the colors of the points), and two plots for weakly supervised metric learning (same) (with red segments for negative constraints and green for positive constraints)}

One of the main advantage of metric-learn is its possibility to be used with

\begin{minted}{python}
from metric_learn import MMC
pairs = np.array([[[1.2, 3.2], [2.3, 5.5]],
                  [[4.5, 2.3], [2.1, 2.3]]])
y_pairs = np.array([1, -1])
mmc = MMC(random_state=42)
mmc.fit(pairs, y_pairs)
mmc.predict(pairs)
\end{minted}

\section{Software Architecture}

\todo[inline]{API: I think we want to insist on this}

\texttt{metric-learn}'s API is designed to be fully compatible with \texttt{scikit-learn} API \cite{scikit-learn}. \texttt{metric-learn} includes different types of algorithms depending on the type of information we want to use to learn the metric. (i) Supervised metric learning algorithms take as an input sample and labels like any supervised learning algorithm, and they will learn a metric that embeds points from the same class closer to each other, and points from a different class further away from each other. (ii) Weakly supervised learning algorithms take as an input tuples of points, and a possible label for the tuple. For example, pairs learners take pairs of points and a label indicating whether the two points are similar (which we want the learned metric to bring closer together) or dissimilar (which we want the learned metric to push away from each other). The API of \texttt{metric-learn} is made so as to allow for the possible development of other types of supervision on tuples. Algorithms that are trained in a supervised way are regular \texttt{Transformers}, just like \texttt{scikit-learn}'s \texttt{LinearDiscriminantAnalysis} for instance. On the other hand, weakly supervised learning algorithms are similar to \texttt{scikit-learn} classifiers, but with input data \texttt{X, y} being tuples and labels of tuples instead of points, i.e. they \texttt{fit} on tuples, that can be represented as 3D arrays, and have a \texttt{predict} method, that gives a prediction for each given tuple. Tuples can be pairs or quadruplets, depending on the specific algorithm used. For instance, \texttt{metric-learn} contains some algorithms that learn on pairs of points and the according labels.

\begin{minted}{python}
from metric_learn import MMC
pairs = np.array([[[1.2, 3.2], [2.3, 5.5]],
                  [[4.5, 2.3], [2.1, 2.3]]])
y_pairs = np.array([1, -1])
mmc = MMC(random_state=42)
mmc.fit(pairs, y_pairs)
mmc.predict(pairs)
\end{minted}

This allows to use out of the box \texttt{scikit-learn}'s scoring functions, and therefore all routines for model selection, including the \texttt{GridSearchCV} object.

\begin{minted}{python}
from sklearn.model_selection import GridSearchCV
grid = {'init': ['random', 'identity', 'covariance']}
model = GridSearchCV(mmc, grid)
model.fit(pairs, y_pairs)
\end{minted}

More details on the 3D array thing:

At the core of \texttt{metric-learn} is the format of input tuples. This format is what allows all these API choices. While a classical input array \texttt{X} in \texttt{scikit-learn} has shape \texttt{(n\_samples, n\_features)}, an array of tuples has shape \texttt{(n\_tuples, t, n\_features)}, where \texttt{t} is the number of elements in a tuple (e.g. 2 for pairs).

Note that \texttt{scikit-learn} accepts arrays in a broad sense, in what they call "array-like" objects. \texttt{metric-learn} too accepts "tuple-like" objects.

Image: a classical array of samples X
an array of pairs of X X[np.array([[1, 2], [3, 4], [2, 1]])]

\todo[inline]{CI and test coverage}

The quality of the code is ensured by an intensive test coverage of 96\%. Every new contribution to the code is automatically checked by a continuous integration platform that ensures a minimal test coverage of the added code, and an minimal variation in test coverage of the overall code.

Contributions: This allows easy contribution for incoming developers, which can have their code automatically checked. What is more, they can instantiate their algorithm as an instance of Metric learner, to benefit from all the methods of it, and help scikit-learn compatibility.

\section{Conclusion}

mention future developments? scalability, more algorithms, etc

scalability: 
- support stochastic solvers
- support sparsity
- form pairs on the fly and do some computations by batch

more algorithms:
- algorithms on triplets
- algorithms for high dimensional data

- toy datasets

improved CI
- pep8

\acks{TODO}

 The quality of comparison to previous (if any) related implementations, w.r.t. run-time, memory requirements, features, to explain that significant progress has been made.

% Manual newpage inserted to improve layout of sample file - not
% needed in general before appendices/bibliography.

\bibliography{metric-learn-jmlr}

\end{document}

### 2019-07-17 generate_figures.py
add figures

@@ -1,0 +1,54 @@
+from metric_learn import ITML, MMC
+import matplotlib.pyplot as plt
+import numpy as np
+
+
+X = np.array([[-3, 2],
+              [-2.9, -0.86],
+              [4.24, 3.08],
+              [4.08, -0.2],
+              [1.58, -3.14],
+              [-0.38, -3.76]])
+
+c = np.array([[0, 4], [1, 2], [3, 5], [2, 3]])
+
+pairs = X[c]
+y_pairs = np.array([1, -1, -1, 1])
+
+def plot_points(pairs, y_pairs, name):
+  x_min, x_max = pairs[:, :, 0].min(), pairs[:, :, 0].max()
+  y_min, y_max = pairs[:, :, 1].min(), pairs[:, :, 1].max()
+
+  x_center = (x_min + x_max)/2
+  y_center = (y_min + y_max)/2
+
+  max_diff = max(x_max - x_min, y_max - y_min)
+  margin = max_diff / 20
+
+  y_lims = (y_center - max_diff/2 - margin, y_center + max_diff/2 + margin)
+  x_lims = (x_center - max_diff/2 - margin, x_center + max_diff/2 + margin)
+
+
+
+  plt.figure()
+
+  for i, edge in enumerate(pairs):
+      plt.plot(edge[:, 0], edge[:, 1], c='green' if y_pairs[i] == 1 else
+      'red', alpha=0.3)
+  plt.scatter(pairs[:, 0, 0], pairs[:, 0, 1], c='b')
+  plt.scatter(pairs[:, 1, 0], pairs[:, 1, 1], c='b')
+  plt.xlim(*x_lims)
+  plt.ylim(*y_lims)
+  plt.axis('equal')
+  plt.savefig(name)
+
+
+plot_points(pairs, y_pairs, 'pairs_without_metric')
+
+mmc = ITML()
+mmc.fit(pairs, y_pairs)
+
+X_e = mmc.transform(X)
+pairs = X_e[c].copy()
+
+plot_points(pairs, y_pairs, 'pairs_with_metric')

### 2019-07-17 metric-learn-jmlr.tex
wip

@@
 \usepackage{minted}
+\usepackage[demo]{graphicx}
+\usepackage{caption}
+\usepackage{subcaption}
 \usepackage[colorinlistoftodos]{todonotes}
@@
 
-\texttt{metric-learn} contains most of the highly cited Mahalanobis metric learning algorithms. Here is a list of the algorithms contained in \texttt{metric-learn}:
-\begin{itemize}
-    \item Supervised Metric Learning algorithms: 
+\texttt{metric-learn} contains most of the highly cited Mahalanobis metric learning algorithms.
+
+\subsection{Mahalanobis metric learning}
+Mahalanobis learning algorithms learn a
+
+Here is a list of the algorithms contained in \texttt{metric-learn}:
+\subsection{Supervised Metric Learning algorithms}
 Supervised algorithms learns a metric that brings closer together points from the same class and spreads further away points from a different class.
@@
 \end{itemize}
-    \item Weakly Supervised Metric Learning algorithms:
+\subsection{Weakly Supervised Metric Learning algorithms}
 Weakly Supervised algorithms use less information than the previous algorihtms: they only take tuple of points and possibly a corresponding label.
 See below for more concrete examples of algorithms learning on tuples.
-    \begin{itemize}
-        \item Pairs Learners: 
+    \subsubsection{Pairs Learners} 
     Pairs learners take as input pair of points, and a corresponding label indicating whether the two points are similar or not.
@@
     \end{itemize}
-        \item Quadruplets Learners: 
+
+    \begin{figure}[H]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.8\linewidth]{pairs_without_metric.png}
+  \caption{Original points, before metric learning}
+  \label{fig:sub1}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.8\linewidth]{pairs_with_metric.png}
+  \caption{Points after metric learning (ITML)}
+  \label{fig:sub2}
+\end{subfigure}
+\caption{Metric learning with pairwise constraints}
+\label{fig:test}
+\end{figure}
+
+    \subsubsection{Quadruplets Learners}
     Quadruplets learners take as input quadruplets of points. They learn a metric that brings the two first points of each quadruplet closer than the two last points are.
@@
         \item LSML: Metric Learning from Relative Comparisons by Minimizing Squared Residual
-        \end{itemize}
-    \end{itemize}
     \end{itemize}

### 2019-07-17 generate_supervised_learning.py
add figures for supervised learning

@@ -1,0 +1,46 @@
+from metric_learn import NCA, MMC
+import matplotlib.pyplot as plt
+import numpy as np
+
+
+X = np.array([[-3, 2],
+              [-2.9, -0.86],
+              [4.24, 3.08],
+              [4.08, -0.2],
+              [1.58, -3.14],
+              [-0.38, -3.76]])
+
+y = np.array([0, 2, 1, 1, 0, 2])
+
+def plot_points(X, y, name):
+  x_min, x_max = X[:, 0].min(), X[:, 0].max()
+  y_min, y_max = X[:, 1].min(), X[:, 1].max()
+
+  x_center = (x_min + x_max)/2
+  y_center = (y_min + y_max)/2
+
+  max_diff = max(x_max - x_min, y_max - y_min)
+  margin = max_diff / 20
+
+  y_lims = (y_center - max_diff/2 - margin, y_center + max_diff/2 + margin)
+  x_lims = (x_center - max_diff/2 - margin, x_center + max_diff/2 + margin)
+
+
+
+  plt.figure()
+
+  plt.scatter(X[:, 0], X[:, 1], c=y)
+  plt.xlim(*x_lims)
+  plt.ylim(*y_lims)
+  plt.axis('equal')
+  plt.savefig(name)
+
+
+plot_points(X, y, 'supervised_without_metric')
+
+nca = NCA()
+nca.fit(X, y)
+
+X_e = nca.transform(X)
+
+plot_points(X_e, y, 'supervised_with_metric')

### 2019-07-26 metric-learn-jmlr.tex
Updates from Overleaf

--- 2019-07-17 metric-learn-jmlr.tex
@@
 \usepackage{minted}
+\usepackage{graphicx}
+\usepackage{caption}
+\usepackage{subcaption}
+\usepackage{stmaryrd}
 \usepackage[colorinlistoftodos]{todonotes}
@@
 
-\jmlrheading{1}{2000}{1-48}{4/00}{10/00}{Marina Meil\u{a} and Michael I. Jordan}
+\jmlrheading{}{}{}{}{}{William de Vazelhes and CJ Carey and Yuan Tang and Nathalie Vauquier and Aur\'elien Bellet}
 
@@
 
-\ShortHeadings{Learning with Mixtures of Trees}{Meil\u{a} and Jordan}
+\ShortHeadings{Metric-learn: Metric Learning in Python}{de Vazelhes, Carey, Tang, Vauquier and Bellet}
 \firstpageno{1}
@@
 
-\author{\name Marina Meil\u{a} \email mmp@stat.washington.edu \\
-       \addr Department of Statistics\\
-       University of Washington\\
-       Seattle, WA 98195-4322, USA
+\author{\name William de Vazelhes \email william.de-vazelhes@inria.fr \\
+       \addr INRIA, France
        \AND
-       \name Michael I.\ Jordan \email jordan@cs.berkeley.edu \\
-       \addr Division of Computer Science and Department of Statistics\\
-       University of California\\
-       Berkeley, CA 94720-1776, USA}
+       \name CJ Carey \email perimosocordiae@gmail.com \\
+       \addr \todo[inline]{TODO}
+       \AND
+       \name Yuan Tang \email terrytangyuan@gmail.com \\
+       \addr \todo[inline]{TODO}
+       \AND
+       \name Nathalie Vauquier \email nathalie.vauquier@inria.fr \\
+       \addr INRIA, France
+       \AND
+       \name Aur\'elien Bellet \email aurelien.bellet@inria.fr \\
+       \addr INRIA, France
+       }
 
-\editor{Leslie Pack Kaelbling}
+\editor{}
 
@@
 \begin{abstract}%   <- trailing '%' for backward compatibility of .sty file
-TODO
+\texttt{metric-learn} is an open source Python package for metric-learning, based on \texttt{numpy}, \texttt{scipy}, and \texttt{scikit-learn}.
+It allows to learn a metric between data points that satisfy desired properties.
+\texttt{metric-learn} stands out from other packages by providing a unified interface for metric learning with several modes of supervision (full supervision or pairs and quadruplets supervision), all being compatible with \texttt{scikit-learn} and allowing to do cross-validation, model selection, and pipelining.
+The package is released under the MIT licence, and the current version (v0.5.0) is downloadable from PyPi.
+The source code is available online at \url{http://github.com/metric-learn/metric-learn}, as well as the documentation at \url{http://metric-learn.github.io/metric-learn/}.
 \end{abstract}
@@
 
-\begin{itemize}
-    \item Metric learning
-    \item Importance of supporting the classic ML workflows for model selection, model evaluation, pipelining with other estimators etc, via sklearn. makes metric-learn stands out from other packages
-    in particular in the case of weakly supervised algorithms
-\end{itemize}
-
+Many approaches in machine learning require a measure of distance between data
+points. Traditionally, practitioners would choose a standard distance metric
+(Euclidean, City-Block, Cosine, etc.) using a priori knowledge of the
+domain. However, it is often difficult to design metrics that are well-suited
+to the particular data and task of interest.
+% \todo[inline]{probably need to add a sentence explaining what is meant by weak supervision}
+Distance metric learning  (or simply, metric learning) (\cite{Bellet13}, \cite{Kulis13}, \cite{Surez18}) aims at
+automatically constructing task-specific distance metrics from data, in a machine learning manner.
+It can deal with various types of supervision: either full supervision, where we know label(s) for each sample, or weaker supervision, where we only have some information about the similarities between samples.
+The learned distance metric can
+then be used to perform various tasks, for instance to retrieve the elements (images, documents) of a database that are semantically closer to a query element.
+It can also be plugged into other machine learning algorithms: popular use-cases include improving the accuracy of nearest neighbors learning models (for classification, regression, anomaly detection...) or biasing the clusters found by clustering algorithms like K-Means towards the intended semantics.
+Finally, metric learning can be used to perform dimensionality reduction.
+These use-cases highlight the importance of integrating metric learning into the rest of the machine learning pipeline and tools.
+% \todo[inline]{quick presentation of the package: implements several algorithms, with sklearn-compatible API for model selection, model evaluation, pipelining with other estimators.
+makes metric-learn stands out from other packages (maybe mention a few of them), in particular in the case of weakly supervised algorithms}
+\texttt{metric-learn} is a package for metric learning in Python, which provides many common metric-learning algorithms unified through a common interface.
+It allows to do metric-learning with different level of supervision.
+% (not only class supervision, unlike most packages are focused on, including another package for metric learning in Python, \texttt{PyDML})
+% , which is one interesting feature allowed by the metric learning paradigm. 
+Its API is compatible with scikit-learn, the leading library for machine learning in Python, allowing for model selection, evaluation, and pipelining, which makes \texttt{metric-learn} stand out from other packages, like \texttt{PyDML} \footnote{\url{https://github.com/jlsuarezdiaz/pyDML}}, which mostly focuses on supervised metric learning algorithms.
 % Acknowledgements should go at the end, before appendices and references
+The rest of this paper is organized as follows.
+We first present an overview of the package, followed by a more detailed description of the available algorithms.
+We then give more details about the API and the software architecture, in particular the compatibility with \texttt{scikit-learn}, and conclude with a discussion of future developments.
 
 \section{Overview of the Package}
-\texttt{metric-learn} is  a
-\todo[inline]{GitHub repo, PyPi, mention current release, license}
-\texttt{metric-learn} only relies on core libraries from the SciPy ecosystem: \texttt{numpy}, \texttt{scipy}, and \texttt{scikit-learn}.
-An optional dependency with \texttt{skggm} (\cite{skggm}) can be installed in order to use a more stable solver when using the LSML algorithm from the package.
-It can be installed from the Python Package Index (PyPI), on python 2.7 and 3.5 or later.
-The source code is available at \url{http://github.com/metric-learn/metric-learn}.
-The development is collaborative and open to all.
+
+\texttt{metric-learn} only relies on core libraries from the SciPy ecosystem: \texttt{numpy}, \texttt{scipy}, and \texttt{scikit-learn}.\footnote{An optional dependency with \texttt{skggm} (\cite{skggm}) can be installed in order to use a more stable solver when using the LSML algorithm from the package.}
+The current release (v.0.5.0) can be installed from the Python Package Index (PyPI), for python 2.7 and 3.5 or later.
+The source code is available on GitHub at \url{http://github.com/metric-learn/metric-learn}.
+The development is collaborative and open to all contributors through the usual GitHub framework (issues and pull requests).
 The code is free of use, provided under the MIT licence.
-The current realease is 0.5.0.
-It includes many refactoring including the ones described in the paper, that improve scikit-learn compatibility.
+The interest of the package from the community is clear, since the project has been starred more than 730 times and forked more than 170 times at the time of writing.
+A detailed documentation, including installation guidelines, the description of the algorithms and the API, as well as example of use done with \texttt{sphinx-gallery}, is available at \url{https://metric-learn.github.io/metric-learn/}.
+The quality of the code is ensured by an intensive test coverage of 96\%.
+Every new contribution to the code is automatically checked by a continuous integration platform that ensures a minimal test coverage of the added code, and an minimal variation in test coverage of the overall code, which allows easy contribution for incoming developers, which can have their code automatically checked.
 
 \section{Available algorithms}
+% \todo[inline]{comparison with other packages ? }
+% Metric learning problems fall into two main categories depending on the type
+% of supervision available about the training data:
 
-\texttt{metric-learn} contains most of the highly cited Mahalanobis metric learning algorithms. Here is a list of the algorithms contained in \texttt{metric-learn}:
-\begin{itemize}
-    \item Supervised Metric Learning algorithms: 
-    Supervised algorithms learns a metric that brings closer together points from the same class and spreads further away points from a different class.
-    \begin{itemize}
-        \item NCA: Neighborhood Components Analysis
-        \item LMNN: Large Margin Nearest Neighbors
-        \item RCA: Relative Components Analysis
-        \item LFDA: Local Fisher Discriminant Analysis
-    \end{itemize}
-    \item Weakly Supervised Metric Learning algorithms:
-    Weakly Supervised algorithms use less information than the previous algorihtms: they only take tuple of points and possibly a corresponding label.
+% \begin{itemize}
+%     \item 
+% - Supervised learning: the algorithm has access to
+%   a set of data points, each of them belonging to a class (label) as in a
+%   standard classification problem.
+%   Broadly speaking, the goal in this setting is to learn a distance metric
+%   that puts points with the same label close together while pushing away
+%   points with different labels.
+% - Weakly supervised learning: the
+%   algorithm has access to a set of data points with supervision only
+%   at the tuple level (typically pairs, triplets, or quadruplets of
+%   data points). A classic example of such weaker supervision is a set of
+%   positive and negative pairs: in this case, the goal is to learn a distance
+%   metric that puts positive pairs close together and negative pairs far away.
+
+% \end{itemize}
+
+\paragraph{General setting} Metric learning problems are generally formulated as an optimization problem where one seeks to find, on the training data $(x_0, ..., x_n)$, the parameters $\theta$ of a distance function $D_\theta(x_i, x_j)$ that optimize some desired objective function $\mathcal{L}_\theta(x_0, ..., x_n) = f(D_\theta(x_0, x_0), ..., D_\theta(x_0, x_n), ..., D_\theta(x_n, x_0), ..., D_\theta(x_n, x_n))$, hoping that this metric will generalize the desired propertie(s) on new incoming test data.
+
+\subparagraph{Mahalanobis metric learning}
+
+In the metric-learn package, all algorithms currently implemented learn so-called Mahalanobis distances.
+Given a real-valued parameter matrix $L$ of shape \texttt{(num\_dims, n\_features)} where \texttt{n\_features} is the
+number of features describing the data, the Mahalanobis distance associated with
+$L$ is defined as follows:
+
+$D(x, x') = \sqrt{(Lx-Lx')^\top(Lx-Lx')}$
+
+In other words, a Mahalanobis distance is a Euclidean distance after a
+linear transformation of the feature space defined by $L$ (taking
+$L$ to be the identity matrix recovers the standard Euclidean distance).
+Mahalanobis distance metric learning can thus be seen as learning a new
+embedding space of dimension \texttt{num\_dims}. Note that when \texttt{num\_dims} is
+smaller than \texttt{n\_features}, this achieves dimensionality reduction.
+
+% Strictly speaking, Mahalanobis distances are "pseudo-metrics": they satisfy
+% three of the properties of a metric (non-negativity, symmetry, triangle inequality) but not
+% necessarily the identity of indiscernibles.
+
+% Mahalanobis distances can also be parameterized by a positive semi-definite 
+% (PSD) matrix $M$:
+
+% $ D(x, x') = \sqrt{(x-x')^\top M(x-x')}$
+
+% Using the fact that a PSD matrix :math:`M` can always be decomposed as
+% $M=L^\top L$ for some  :math:`L`, one can show that both
+% parameterizations are equivalent. In practice, an algorithm may thus solve
+% the metric learning problem with respect to either :math:`M` or :math:`L`.
+
+\paragraph{Supervised Learners}
+The first possibility to learn this metric is from class supervision.
+We can learn a metric that brings closer together points from the same class and spreads further away points from a different class.
+For instance, in figure \ref{fig:full}, the class is the identity of a person of an image, and we want to put closer together images from the same person.Currently, \texttt{metric-learn} contains the following algorithms for full supervision: Neighborhood Components Analysis (NCA) (\cite{Goldberger04}), Large Margin Nearest Neighbors (LMNN) (\cite{Weinberger09}), Relative Components Analysis (RCA) (\cite{Shental02}), Local Fisher Discriminant Analysis (LFDA) (\cite{Sugiyama07}) and Metric Learning for Kernel Regression (MLKR) (\cite{Weinberger07}) (which uses a continuous label like in regressions, instead of a class label).
+
+% \todo[inline]{ces figures ne sont pas très vendeuses.
+je propose de faire une figure conceptuelle en prenant l'exemple des faces expliquant la différence entre le supervisé, les paires et les quadruplets.
+on peut s'inspirer de \url{http://cedric.cnam.fr/~thomen/papers/Law_ICCV_2013_QWise.pdf}}
+
+% \begin{figure}[H]
+% \centering
+% \begin{subfigure}{.5\textwidth}
+%   \centering
+%   \includegraphics[width=.8\linewidth]{supervised_without_metric.png}
+%   \caption{Original points, before metric learning}
+%   \label{fig:sub1}
+% \end{subfigure}%
+% \begin{subfigure}{.5\textwidth}
+%   \centering
+%   \includegraphics[width=.8\linewidth]{supervised_with_metric.png}
+%   \caption{points after metric learning (ITML)}
+%   \label{fig:sub2}
+% \end{subfigure}
+% \caption{Metric learning with pairwise constraints}
+% \label{fig:test}
+% \end{figure}
+
+% \paragraph{Weakly Supervised Metric Learning algorithms}
+% Weakly Supervised algorithms use less information than the previous algorithms: they only take tuple of points and possibly a corresponding label.
 See below for more concrete examples of algorithms learning on tuples.
-    \begin{itemize}
-        \item Pairs Learners: 
-        Pairs learners take as input pair of points, and a corresponding label indicating whether the two points are similar or not.
-They will learn a metric that brings points from a similar pair closer together, and points from a dissimilar pair further away from each other.
-        \begin{itemize}
-            \item MMC: Mahalanobis Metric for Clustering
-            \item SDML: Sparse High-Dimensional Metric Learning (SDML)
-            \item ITML: Information Theoretic Metric Learning (ITML)
-        \end{itemize}
-        \item Quadruplets Learners: 
-        Quadruplets learners take as input quadruplets of points. They learn a metric that brings the two first points of each quadruplet closer than the two last points are.
-        \begin{itemize}
-            \item LSML: Metric Learning from Relative Comparisons by Minimizing Squared Residual
-        \end{itemize}
-    \end{itemize}
-\end{itemize}
+\subparagraph{Pairs Learners} 
+Another way is to have pair of points as inputs, and a corresponding label indicating whether the two points are similar or not.
+We can then learn a metric that brings points from a similar pair closer together, and points from a dissimilar pair further away from each other.
+For instance, we could want to put closer together in the embedding space images from the same person (i.e.
+similar), and further away images from a different person (i.e.
+dissimilar) (see figure \ref{fig:pairs}).
+Currently, \texttt{metric-learn} contains the following algorithms for pairs supervision: Mahalanobis Metric for Clustering (MMC) (\cite{Xiang08}), Sparse High-Dimensional Metric Learning (SDML) (\cite{Qi09}), Information Theoretic Metric Learning (ITML) (\cite{Davis07}).
 
-\todo[inline]{Mention and link documentation}
+% \begin{figure}[H]
+% \centering
+% \begin{subfigure}{.5\textwidth}
+%   \centering
+%   \includegraphics[width=.8\linewidth]{pairs_without_metric.png}
+%   \caption{Original points, before metric learning}
+%   \label{fig:sub1}
+% \end{subfigure}%
+% \begin{subfigure}{.5\textwidth}
+%   \centering
+%   \includegraphics[width=.8\linewidth]{pairs_with_metric.png}
+%   \caption{points after metric learning (ITML)}
+%   \label{fig:sub2}
+% \end{subfigure}
+% \caption{Metric learning with pairwise constraints}
+% \label{fig:test}
+% \end{figure}
 
-\textbf{Documentation} A detailed documentation, including installation guidelines, the description of the algorithms and the API, as well as example of use done with \texttt{sphinx-gallery}, is available at \url{https://metric-learn.github.io/metric-learn/}.
-It also includes \texttt{intersphinx} bindings for more efficient navigation through the documentation of the whole \texttt{scipy} ecoystem.
+\subparagraph{Quadruplets Learners}
+The last possibility to learn a metric in \texttt{metric-learn} currently is to take as input quadruplets of points.
+We will learn a metric that brings the two first points of each quadruplet closer than the two last points are.
+This can be used for instance to learn a metric space in which closer points have closer values of a given attributes.
+In figure \ref{fig:quadruplets} for instance, we want to learn a metric that puts closer together people from approximately the same age.
+Currently the only algorithm in \texttt{metric-learn} for learning on quadruplets is Metric Learning from Relative Comparisons by Minimizing Squared Residual (LSML) (\cite{Liu12}).
 
-\todo[inline]{Code example to showcase usage, potentially with compelling visualization?}
+\begin{figure}[H]
+  \centering
+  \begin{subfigure}[t]{0.3\textwidth}
+     \centering 
+     \begin{tabular}{c|c}
+     \texttt{X} & \texttt{y} \\ \hline \\
+   \includegraphics[scale=0.15]{Tom_Cruise_0010.jpg} & 'Tom' \\
+   \includegraphics[scale=0.15]{Kate_Winslet_0001.jpg} & 'Kate' \\
+\end{tabular}
+     \caption{}
+     \label{fig:full}
+     \end{subfigure}
+  \hspace{5pt}
+  \begin{subfigure}[t]{0.3\textwidth}
+     \centering
+     \begin{tabular}{c|c}
+     \texttt{pairs} & \texttt{y\_pairs} \\  \hline \\
+\includegraphics[scale=0.15]{Aaron_Peirsol_0001.jpg} \includegraphics[scale=0.15]{Aaron_Peirsol_0002.jpg} & 1\\  \includegraphics[scale=0.15]{Angelina_Jolie_0001.jpg}   \includegraphics[scale=0.15]{Kate_Winslet_0001.jpg} & -1 \\
+\end{tabular}
+     \caption{}
+     \label{fig:pairs}
+  \end{subfigure}
+  \hspace{5pt}
+  \begin{subfigure}[t]{0.3\textwidth}
+     \centering
 
-\todo[inline]{what about four plots: two plots for supervised metric learning: before and after learning the metric (classes are the colors of the points), and two plots for weakly supervised metric learning (same) (with red segments for negative constraints and green for positive constraints)}
+          \begin{tabular}{c}
+     \texttt{quadruplets} \\ \hline \\
+     \includegraphics[scale=0.15]{Kate_Winslet_0001.jpg} \includegraphics[scale=0.15]{Nicole_Kidman_0001.jpg} \includegraphics[scale=0.15]{Daniel_Radcliffe_0001.jpg} \includegraphics[scale=0.15]{Kirk_Douglas_0001.jpg} \\
+          \includegraphics[scale=0.15]{Leonardo_DiCaprio_0001.jpg} \includegraphics[scale=0.15]{Angelina_Jolie_0001.jpg} \includegraphics[scale=0.15]{Daryl_Sabara_0001.jpg} \includegraphics[scale=0.15]{Carl_Reiner_0002.jpg} \\
+\end{tabular}
 
-One of the main advantage of metric-learn is its possibility to be used with
+     \caption{}
+     \label{fig:quadruplets}
+  \end{subfigure}
+  \caption{Different types of supervision.
+\textbf{\ref{fig:full} Full supervision:} we want samples from the same label to be closer than samples from different labels.
+\textbf{\ref{fig:pairs} Pairs supervision:} we want similar samples (1) (the same person) to be closer than dissimilar samples (-1) (different persons).
+\textbf{\ref{fig:quadruplets} Quadruplets supervision:} no labels.
+We want the two leftmost samples (not much age difference) to be closer than the two rightmost samples (bigger age difference).}
+  \label{fig:contour}
+\end{figure}
+
+\section{Software Architecture and API}
+
+\texttt{metric-learn}'s goal is to provide a unified interface to all metric-learning algorithms, including the ones that have an input format that differs from classical machine learning ones, and allow to combine them with scikit-learn objects.
+As such, all metric-learners from \texttt{metric-learn} inherit from an abstract \texttt{BaseMetricLearner} object.
+% \todo[inline]{Should I talk about the classes etc ? Not useful since they are not visible by the user right ?}
+% \todo[inline]{I think it can be nice, this is not documentation, but explaining how things are coded}
+
+\texttt{metric-learn}'s API is designed to be fully compatible with \texttt{scikit-learn} API (\cite{scikit-learn}).
+
+\paragraph{Mahalanobis metric learning}
+
+Since they learn a Mahalanobis matrix, all algorithm also inherit from the \texttt{MahalanobisMixin} interface.
+This latter can return the mahalanobis matrix with the method \texttt{get\_mahalanobis\_matrix}.
+Like scikit-learn's algorithms, it can also return a \texttt{components\_} attributes that is the transformation matrix \texttt{L}.
+Since all these metric learner learn this implicit transformation $L$, they can project the input in the new space using \texttt{transform}.
+A distance function can also be extracted from them using \texttt{get\_metric}, that can further be used in any scikit-learn estimator like \texttt{KMeansClustering}.
+They can also return the distance between a dataset of pairs (in the 3D array format (see below), using \texttt{score\_pairs}.
+
+\paragraph{Supervised metric learners}
+
+Supervised metric learners inherit from scikit-learn base class \texttt{TransformerMixin}, just like \texttt{scikit-learn}'s \texttt{LinearDiscriminantAnalysis} for instance, and they behave as such.
+% (they pass all scikit-learn's "estimators checks"). 
+%  They can easily be pipelined with nearest neighbors estimators to improve their performance: \texttt{better\_knn = sklearn.pipeline.make\_pipeline(NCA(), KNeighborsClassifier())} (TODO: sth to explain NCA is in metric-learn)
+
+\paragraph{Weakly Supervised metric learners} 
+% Algorithms that learn on pairs or quadruplets take as input 3D array of shape \texttt{(n\_samples, t, n\_features)} (where...).
+In other words, sampling one element along the first dimension returns one tuple (e.g.
+a pair), that is a matrix of t lines corresponding to the t elements in the tuple.
+% This format is what allows all these API choices, since slicing along the 1st dimension
+% While a classical input array \texttt{X} in \texttt{scikit-learn} has shape \texttt{(n\_samples, n\_features)}, an array of tuples has shape \texttt{(n\_tuples, t, n\_features)}, where \texttt{t} is the number of elements in a tuple (e.g.
+2 for pairs).
+% Note that \texttt{scikit-learn} accepts arrays in a broad sense, in what they call "array-like" objects. \texttt{metric-learn} too accepts "tuple-like" objects.
+Weakly supervised learning algorithms are similar to \texttt{scikit-learn} classifiers, but they \texttt{fit} on tuples that can be given as 3D arrays, instead of 2D arrays-like of points \texttt{X}.
+They have a \texttt{predict} method, that gives a prediction for each given tuple (e.g.
+for pairs it predicts whether the two points in a pair or similar or not).
+Tuples can be pairs or quadruplets, depending on the specific algorithm used.
+Below is a code example for pairs supervision:
 
 \begin{minted}{python}
+import numpy as np
 from metric_learn import MMC
-pairs = np.array([[[1.2, 3.2], [2.3, 5.5]],
-                  [[4.5, 2.3], [2.1, 2.3]]])
+pairs = np.array([[[1.2, 3.2], [2.3, 5.5]], [[4.5, 2.3], [2.1, 2.3]]])
 y_pairs = np.array([1, -1])
-mmc = MMC(random_state=42)
+mmc = MMC()
 mmc.fit(pairs, y_pairs)
-mmc.predict(pairs)
 \end{minted}
-
-\section{Software Architecture}
-
-\todo[inline]{API: I think we want to insist on this}
-
-\texttt{metric-learn}'s API is designed to be fully compatible with \texttt{scikit-learn} API \cite{scikit-learn}.
-\texttt{metric-learn} includes different types of algorithms depending on the type of information we want to use to learn the metric.
-(i) Supervised metric learning algorithms take as an input sample and labels like any supervised learning algorithm, and they will learn a metric that embeds points from the same class closer to each other, and points from a different class further away from each other.
-(ii) Weakly supervised learning algorithms take as an input tuples of points, and a possible label for the tuple.
-For example, pairs learners take pairs of points and a label indicating whether the two points are similar (which we want the learned metric to bring closer together) or dissimilar (which we want the learned metric to push away from each other).
-The API of \texttt{metric-learn} is made so as to allow for the possible development of other types of supervision on tuples.
-Algorithms that are trained in a supervised way are regular \texttt{Transformers}, just like \texttt{scikit-learn}'s \texttt{LinearDiscriminantAnalysis} for instance.
-On the other hand, weakly supervised learning algorithms are similar to \texttt{scikit-learn} classifiers, but with input data \texttt{X, y} being tuples and labels of tuples instead of points, i.e.
-they \texttt{fit} on tuples, that can be represented as 3D arrays, and have a \texttt{predict} method, that gives a prediction for each given tuple.
-Tuples can be pairs or quadruplets, depending on the specific algorithm used.
-For instance, \texttt{metric-learn} contains some algorithms that learn on pairs of points and the according labels.
-
-\begin{minted}{python}
-from metric_learn import MMC
-pairs = np.array([[[1.2, 3.2], [2.3, 5.5]],
-                  [[4.5, 2.3], [2.1, 2.3]]])
-y_pairs = np.array([1, -1])
-mmc = MMC(random_state=42)
-mmc.fit(pairs, y_pairs)
-mmc.predict(pairs)
-\end{minted}
+% mmc.predict(pairs)
 
 This allows to use out of the box \texttt{scikit-learn}'s scoring functions, and therefore all routines for model selection, including the \texttt{GridSearchCV} object.
+Also note that for every weakly supervised learner, there is a \texttt{\_Supervised} version (e.g.
+for \texttt{MMC} there is \texttt{MMC\_Supervised}, that is a supervised version that samples pairs of similar samples from intra-class samples, and dissimilar samples from inter-class samples, and fits \texttt{MMC} on these pairs).
 
-\begin{minted}{python}
-from sklearn.model_selection import GridSearchCV
-grid = {'init': ['random', 'identity', 'covariance']}
-model = GridSearchCV(mmc, grid)
-model.fit(pairs, y_pairs)
-\end{minted}
+% \begin{minted}{python}
+% from sklearn.model_selection import GridSearchCV
+% model = GridSearchCV(mmc, {'init': ['random', 'identity', 'covariance']})
+% model.fit(pairs, y_pairs)
+% \end{minted}
 
-More details on the 3D array thing:
+\section{Conclusion and Future Work}
 
-At the core of \texttt{metric-learn} is the format of input tuples.
-This format is what allows all these API choices.
-While a classical input array \texttt{X} in \texttt{scikit-learn} has shape \texttt{(n\_samples, n\_features)}, an array of tuples has shape \texttt{(n\_tuples, t, n\_features)}, where \texttt{t} is the number of elements in a tuple (e.g.
-2 for pairs).
+In conclusion, we presented a scikit-learn compatible package for metric-learning.
+This package is under active development and many improvements are planned to be added to the package.
+First, the scalability of the package could be improved by supporting stochastic solvers (like stochastic gradient descent), and/or to form pairs on the fly and do computations by batch, in order not to load every sample in memory at once.
+We could also take advantage of possible sparse input format to reduce computations.
+More algorithms could also be added to the package, like algorithms learning on triplets, algorithms that can deal with high dimensional data.
+We could also allow metric learn to be used when having multiple labels for each sample, and allow different kind of supervisions, like semi-supervised learning for instance.
+We could also improve the ease of use by adding more examples and adding toy datasets to play with quickly and see the interest of the weaker forms of supervision.
+About the overall quality of the software, we intend to improve the continuous integration routines, for instance by ensuring a PEP8 syntax checker is run on every pull request, a continuous integration for the documentation.
 
-Note that \texttt{scikit-learn} accepts arrays in a broad sense, in what they call "array-like" objects. \texttt{metric-learn} too accepts "tuple-like" objects.
+\acks
 
-Image: a classical array of samples X
-an array of pairs of X X[np.array([[1, 2], [3, 4], [2, 1]])]
+We would like to be particularly thankful to Inria for funding two years of development of the package.
+Also we would like to thanks \texttt{scikit-learn} developers for fruitful discussions, assistance for developing the new API, and funding for conference attending.
+\todo[inline]{reformulate + other acknowledgments}
 
-\todo[inline]{CI and test coverage}
-
-The quality of the code is ensured by an intensive test coverage of 96\%.
-Every new contribution to the code is automatically checked by a continuous integration platform that ensures a minimal test coverage of the added code, and an minimal variation in test coverage of the overall code.
-
-Contributions: This allows easy contribution for incoming developers, which can have their code automatically checked.
-What is more, they can instantiate their algorithm as an instance of Metric learner, to benefit from all the methods of it, and help scikit-learn compatibility.
-
-\section{Conclusion}
-
-mention future developments? scalability, more algorithms, etc
-
-scalability: 
-- support stochastic solvers
-- support sparsity
-- form pairs on the fly and do some computations by batch
-
-more algorithms:
-- algorithms on triplets
-- algorithms for high dimensional data
-
-- toy datasets
-
-improved CI
-- pep8
-
-\acks{TODO}
-
- The quality of comparison to previous (if any) related implementations, w.r.t. run-time, memory requirements, features, to explain that significant progress has been made.
+%  The quality of comparison to previous (if any) related implementations, w.r.t. run-time, memory requirements, features, to explain that significant progress has been made.
 

### 2019-07-29 metric-learn-jmlr.tex
Test reverting Overleaf change

@@
 We can then learn a metric that brings points from a similar pair closer together, and points from a dissimilar pair further away from each other.
+In practice, this is a much easier type of supervised task to crowd source in particular when there are many labels, since agents will only need to say whether images are similar or not, and will not have to assign a label in a huge list of those.
 For instance, we could want to put closer together in the embedding space images from the same person (i.e.
@@
 This can be used for instance to learn a metric space in which closer points have closer values of a given attributes.
+This is very useful if the attribute is not binary (e.g.
+same person/not same person), like in pairs learning, but is for example continuous (e.g.
+similar age, very different age, not much different age).
 In figure \ref{fig:quadruplets} for instance, we want to learn a metric that puts closer together people from approximately the same age.
@@
      \texttt{quadruplets} \\ \hline \\
-     \includegraphics[scale=0.15]{Kate_Winslet_0001.jpg} \includegraphics[scale=0.15]{Nicole_Kidman_0001.jpg} \includegraphics[scale=0.15]{Daniel_Radcliffe_0001.jpg} \includegraphics[scale=0.15]{Kirk_Douglas_0001.jpg} \\
-          \includegraphics[scale=0.15]{Leonardo_DiCaprio_0001.jpg} \includegraphics[scale=0.15]{Angelina_Jolie_0001.jpg} \includegraphics[scale=0.15]{Daryl_Sabara_0001.jpg} \includegraphics[scale=0.15]{Carl_Reiner_0002.jpg} \\
+     D(\includegraphics[scale=0.15]{Kate_Winslet_0001.jpg}, \includegraphics[scale=0.15]{Nicole_Kidman_0001.jpg}) < D\includegraphics[scale=0.15]{Daniel_Radcliffe_0001.jpg}, \includegraphics[scale=0.15]{Kirk_Douglas_0001.jpg}) \\
+        D(  \includegraphics[scale=0.15]{Leonardo_DiCaprio_0001.jpg}, \includegraphics[scale=0.15]{Angelina_Jolie_0001.jpg}) <   D(\includegraphics[scale=0.15]{Daryl_Sabara_0001.jpg},  \includegraphics[scale=0.15]{Carl_Reiner_0002.jpg}) \\
 \end{tabular}
@@
 Also we would like to thanks \texttt{scikit-learn} developers for fruitful discussions, assistance for developing the new API, and funding for conference attending.
-\todo[inline]{reformulate + other acknowledgments}
+\todo[inline]{reformulate + "'(t    other acknowledgments}
 

### 2019-08-07 metric-learn-jmlr.tex
Add back affiliation
Updates from Overleaf

--- 2019-07-26 metric-learn-jmlr.tex
@@
 \usepackage{minted}
+\usepackage[utf8]{inputenc} % allow utf-8 input
 \usepackage{graphicx}
+\usepackage{svg}
 \usepackage{caption}
@@
 \usepackage[colorinlistoftodos]{todonotes}
+\newcommand{\aurelien}[1]{\todo[inline,caption={},color=orange!40]{{\it Aurelien:~}#1}}
+\newcommand{\william}[1]{\todo[inline,caption={},color=blue!40]{{\it William:~}#1}}
+%\renewcommand{\aurelien}[1]{}
+
+\graphicspath{{images/}}
 
@@
 
-\ShortHeadings{Metric-learn: Metric Learning in Python}{de Vazelhes, Carey, Tang, Vauquier and Bellet}
+\ShortHeadings{metric-learn: Metric Learning Algorithms in Python}{de Vazelhes, Carey, Tang, Vauquier and Bellet}
 \firstpageno{1}
@@
 
-\title{Metric-learn: Metric Learning in Python}
+\title{metric-learn: Metric Learning Algorithms in Python}
 
@@
        \name CJ Carey \email perimosocordiae@gmail.com \\
-       \addr \todo[inline]{TODO}
+       \addr \aurelien{CJ must put in his affiliation}
        \AND
        \name Yuan Tang \email terrytangyuan@gmail.com \\
-       \addr \todo[inline]{TODO}
+       \addr Ant Financial, United States
        \AND
@@
 \begin{abstract}%   <- trailing '%' for backward compatibility of .sty file
-\texttt{metric-learn} is an open source Python package for metric-learning, based on \texttt{numpy}, \texttt{scipy}, and \texttt{scikit-learn}.
-It allows to learn a metric between data points that satisfy desired properties.
-\texttt{metric-learn} stands out from other packages by providing a unified interface for metric learning with several modes of supervision (full supervision or pairs and quadruplets supervision), all being compatible with \texttt{scikit-learn} and allowing to do cross-validation, model selection, and pipelining.
-The package is released under the MIT licence, and the current version (v0.5.0) is downloadable from PyPi.
-The source code is available online at \url{http://github.com/metric-learn/metric-learn}, as well as the documentation at \url{http://metric-learn.github.io/metric-learn/}.
+\texttt{metric-learn} is an open source Python package implementing supervised and weakly-supervised distance metric learning algorithms.
+As part of \texttt{scikit-learn-contrib}, it provides a unified interface compatible with \texttt{scikit-learn} which allows to easily perform cross-validation, model selection, and pipelining with other machine learning estimators.
+\texttt{metric-learn} is thoroughly tested and available on PyPi under the MIT licence.
+%The source code is available online at \url{http://github.com/scikit-learn-contrib/metric-learn}, as well as the documentation at \url{http://metric-learn.github.io/metric-learn/}.
 \end{abstract}
@@
 % \todo[inline]{probably need to add a sentence explaining what is meant by weak supervision}
-Distance metric learning  (or simply, metric learning) (\cite{Bellet13}, \cite{Kulis13}, \cite{Surez18}) aims at
+Distance metric learning, or simply metric learning \citep{Bellet15,Kulis13}, aims at
 automatically constructing task-specific distance metrics from data, in a machine learning manner.
-It can deal with various types of supervision: either full supervision, where we know label(s) for each sample, or weaker supervision, where we only have some information about the similarities between samples.
+A key advantage of metric learning is that it can be applied beyond the standard supervised learning setting (data points associated with labels), in situations where only weaker forms of supervision are available (e.g., pairs of points that should be similar/dissimilar).
 The learned distance metric can
@@
 makes metric-learn stands out from other packages (maybe mention a few of them), in particular in the case of weakly supervised algorithms}
-\texttt{metric-learn} is a package for metric learning in Python, which provides many common metric-learning algorithms unified through a common interface.
-It allows to do metric-learning with different level of supervision.
+
+\texttt{metric-learn} is an open source package for metric learning in Python, which implements many popular metric-learning algorithms with different levels of supervision through a unified interface.
+%\footnote{This makes \texttt{metric-learn} stand out from other packages, like \texttt{PyDML} (\url{https://github.com/jlsuarezdiaz/pyDML}) which focuses on supervised metric learning algorithms.}
 % (not only class supervision, unlike most packages are focused on, including another package for metric learning in Python, \texttt{PyDML})
 % , which is one interesting feature allowed by the metric learning paradigm. 
-Its API is compatible with scikit-learn, the leading library for machine learning in Python, allowing for model selection, evaluation, and pipelining, which makes \texttt{metric-learn} stand out from other packages, like \texttt{PyDML} \footnote{\url{https://github.com/jlsuarezdiaz/pyDML}}, which mostly focuses on supervised metric learning algorithms.
+Its API is compatible with scikit-learn, the leading library for machine learning in Python, allowing for streamlined model selection, evaluation, and pipelining with other estimators.
+The project is developed collaboratively through GitHub, its code is thoroughly tested using continuous integration tools.
+It was recently included to \texttt{scikit-learn-contrib}, which hosts high-quality \texttt{scikit-learn}-compatible projects.\footnote{\url{https://github.com/scikit-learn-contrib/scikit-learn-contrib}}
+\aurelien{decide if we include or not a few sentences on positioning wrt other packages}
+
+%, which makes \texttt{metric-learn} stand out from other packages, like \texttt{PyDML} \footnote{\url{https://github.com/jlsuarezdiaz/pyDML}}, which mostly focuses on supervised metric learning algorithms.
 % Acknowledgements should go at the end, before appendices and references
-The rest of this paper is organized as follows.
+%The rest of this paper is organized as follows.
 We first present an overview of the package, followed by a more detailed description of the available algorithms.
@@
 
-\section{Overview of the Package}
+% \aurelien{we could put the presentation of metric learning and supervision as Sec 2, and put the implemented algorithms in the overview of the package section as Section 3}
 
-\texttt{metric-learn} only relies on core libraries from the SciPy ecosystem: \texttt{numpy}, \texttt{scipy}, and \texttt{scikit-learn}.\footnote{An optional dependency with \texttt{skggm} (\cite{skggm}) can be installed in order to use a more stable solver when using the LSML algorithm from the package.}
-The current release (v.0.5.0) can be installed from the Python Package Index (PyPI), for python 2.7 and 3.5 or later.
-The source code is available on GitHub at \url{http://github.com/metric-learn/metric-learn}.
-The development is collaborative and open to all contributors through the usual GitHub framework (issues and pull requests).
-The code is free of use, provided under the MIT licence.
-The interest of the package from the community is clear, since the project has been starred more than 730 times and forked more than 170 times at the time of writing.
-A detailed documentation, including installation guidelines, the description of the algorithms and the API, as well as example of use done with \texttt{sphinx-gallery}, is available at \url{https://metric-learn.github.io/metric-learn/}.
-The quality of the code is ensured by an intensive test coverage of 96\%.
-Every new contribution to the code is automatically checked by a continuous integration platform that ensures a minimal test coverage of the added code, and an minimal variation in test coverage of the overall code, which allows easy contribution for incoming developers, which can have their code automatically checked.
+\section{Background on Metric Learning} \label{metriclearning}
 
-\section{Available algorithms}
 % \todo[inline]{comparison with other packages ? }
@@
 
-\paragraph{General setting} Metric learning problems are generally formulated as an optimization problem where one seeks to find, on the training data $(x_0, ..., x_n)$, the parameters $\theta$ of a distance function $D_\theta(x_i, x_j)$ that optimize some desired objective function $\mathcal{L}_\theta(x_0, ..., x_n) = f(D_\theta(x_0, x_0), ..., D_\theta(x_0, x_n), ..., D_\theta(x_n, x_0), ..., D_\theta(x_n, x_n))$, hoping that this metric will generalize the desired propertie(s) on new incoming test data.
+% \paragraph{General setting} Metric learning problems are generally formulated as an optimization problem where one seeks to find, on the training data $(x_0, ..., x_n)$, the parameters $\theta$ of a distance function $D_\theta(x_i, x_j)$ that optimize some desired objective function $\mathcal{L}_\theta(x_0, ..., x_n) = f(D_\theta(x_0, x_0), ..., D_\theta(x_0, x_n), ..., D_\theta(x_n, x_0), ..., D_\theta(x_n, x_n))$, hoping that this metric will generalize the desired propertie(s) on new incoming test data.
 
-\subparagraph{Mahalanobis metric learning}
-
-In the metric-learn package, all algorithms currently implemented learn so-called Mahalanobis distances.
-Given a real-valued parameter matrix $L$ of shape \texttt{(num\_dims, n\_features)} where \texttt{n\_features} is the
-number of features describing the data, the Mahalanobis distance associated with
-$L$ is defined as follows:
-
-$D(x, x') = \sqrt{(Lx-Lx')^\top(Lx-Lx')}$
-
-In other words, a Mahalanobis distance is a Euclidean distance after a
-linear transformation of the feature space defined by $L$ (taking
-$L$ to be the identity matrix recovers the standard Euclidean distance).
+Metric learning is generally formulated as an optimization problem where one seeks to find the parameters of a distance function that minimize some objective function over the input data.
+All algorithms currently implemented in \texttt{metric-learn} learn so-called Mahalanobis distances.
+Given a real-valued parameter matrix $L$ of shape \texttt{(n\_components, n\_features)} where \texttt{n\_features} is the
+number of features describing the data, the associated Mahalanobis distance between two points $x$ and $x'$ is defined as $D_L(x, x') = \sqrt{(Lx-Lx')^\top(Lx-Lx')}$.
+This is equivalent to a Euclidean distance after the linear transformation of the feature space defined by $L$ (taking $L$ to be the identity matrix recovers the standard Euclidean distance).
 Mahalanobis distance metric learning can thus be seen as learning a new
-embedding space of dimension \texttt{num\_dims}. Note that when \texttt{num\_dims} is
-smaller than \texttt{n\_features}, this achieves dimensionality reduction.
+embedding space of (potentially reduced) dimension \texttt{n\_components}.
+Note that $D_L$ can also be written as $D_L(x, x') = \sqrt{(x - x')^\top M (x - x')}$, where $M = L^\top L$ is called the Mahalanobis matrix.
 
@@
 
-\paragraph{Supervised Learners}
-The first possibility to learn this metric is from class supervision.
-We can learn a metric that brings closer together points from the same class and spreads further away points from a different class.
-For instance, in figure \ref{fig:full}, the class is the identity of a person of an image, and we want to put closer together images from the same person.Currently, \texttt{metric-learn} contains the following algorithms for full supervision: Neighborhood Components Analysis (NCA) (\cite{Goldberger04}), Large Margin Nearest Neighbors (LMNN) (\cite{Weinberger09}), Relative Components Analysis (RCA) (\cite{Shental02}), Local Fisher Discriminant Analysis (LFDA) (\cite{Sugiyama07}) and Metric Learning for Kernel Regression (MLKR) (\cite{Weinberger07}) (which uses a continuous label like in regressions, instead of a class label).
-
+Metric learning algorithms can be classified according to the form of data supervision that they need to learn the metric.
+\texttt{metric-learn} currently implements algorithms that fall into the following categories:
+\begin{itemize}
+\item \textbf{Supervised learners} learn from a standard labeled dataset and essentially aim to bring closer together points from the same class while spreading further away points from a different class.
+For instance, the class may be the identity of a person in an image (see Figure~\ref{fig:full}).
 % \todo[inline]{ces figures ne sont pas très vendeuses.
@@
 on peut s'inspirer de \url{http://cedric.cnam.fr/~thomen/papers/Law_ICCV_2013_QWise.pdf}}
-
 % \begin{figure}[H]
@@
 % \end{figure}
-
 % \paragraph{Weakly Supervised Metric Learning algorithms}
@@
 See below for more concrete examples of algorithms learning on tuples.
-\subparagraph{Pairs Learners} 
+\item \textbf{Pair learners.} 
 Another way is to have pair of points as inputs, and a corresponding label indicating whether the two points are similar or not.
 We can then learn a metric that brings points from a similar pair closer together, and points from a dissimilar pair further away from each other.
-For instance, we could want to put closer together in the embedding space images from the same person (i.e.
-similar), and further away images from a different person (i.e.
-dissimilar) (see figure \ref{fig:pairs}).
-Currently, \texttt{metric-learn} contains the following algorithms for pairs supervision: Mahalanobis Metric for Clustering (MMC) (\cite{Xiang08}), Sparse High-Dimensional Metric Learning (SDML) (\cite{Qi09}), Information Theoretic Metric Learning (ITML) (\cite{Davis07}).
-
+In many applications, this can be a much easier type of supervision to collect, in particular when there are many labels.
+%, since annotators will only need to say whether images are similar or not, and will not have to assign a label in a huge list of those.
+For instance, it is much easier for an annotator to decide whether two images correspond to the same person than to associate an individual image to the right identity among a large number of people (see Figure~\ref{fig:pairs}).
 % \begin{figure}[H]
@@
 % \end{figure}
-
-\subparagraph{Quadruplets Learners}
-The last possibility to learn a metric in \texttt{metric-learn} currently is to take as input quadruplets of points.
-We will learn a metric that brings the two first points of each quadruplet closer than the two last points are.
+\item \textbf{Quadruplet Learners.}
+The last possibility to learn a metric in \texttt{metric-learn} is to take as input quadruplets of points.
+We will learn a metric that brings the two first points of each quadruplet closer than the two last points.
 This can be used for instance to learn a metric space in which closer points have closer values of a given attributes.
-In figure \ref{fig:quadruplets} for instance, we want to learn a metric that puts closer together people from approximately the same age.
-Currently the only algorithm in \texttt{metric-learn} for learning on quadruplets is Metric Learning from Relative Comparisons by Minimizing Squared Residual (LSML) (\cite{Liu12}).
+This is very useful if the attribute is not binary like in pair supervision (e.g.
+same person/not same person), but is rather continuous and difficult to annotate precisely (e.g., the age of a person on an image, see Figure \ref{fig:quadruplets}).
+%for instance, we want to learn a metric that puts closer together people from approximately the same age.
+\end{itemize}
 
 \begin{figure}[H]
-  \centering
-  \begin{subfigure}[t]{0.3\textwidth}
-     \centering 
-     \begin{tabular}{c|c}
-     \texttt{X} & \texttt{y} \\ \hline \\
-   \includegraphics[scale=0.15]{Tom_Cruise_0010.jpg} & 'Tom' \\
-   \includegraphics[scale=0.15]{Kate_Winslet_0001.jpg} & 'Kate' \\
-\end{tabular}
-     \caption{}
+  \begin{subfigure}
+     \includegraphics[scale=0.2]{labels.pdf}
      \label{fig:full}
 \end{subfigure}
-  \hspace{5pt}
-  \begin{subfigure}[t]{0.3\textwidth}
-     \centering
-     \begin{tabular}{c|c}
-     \texttt{pairs} & \texttt{y\_pairs} \\  \hline \\
-\includegraphics[scale=0.15]{Aaron_Peirsol_0001.jpg} \includegraphics[scale=0.15]{Aaron_Peirsol_0002.jpg} & 1\\  \includegraphics[scale=0.15]{Angelina_Jolie_0001.jpg}   \includegraphics[scale=0.15]{Kate_Winslet_0001.jpg} & -1 \\
-\end{tabular}
-     \caption{}
+ \begin{subfigure}
+     \includegraphics[scale=0.2]{pairs.pdf}
      \label{fig:pairs}
   \end{subfigure}
-  \hspace{5pt}
-  \begin{subfigure}[t]{0.3\textwidth}
-     \centering
-
-          \begin{tabular}{c}
-     \texttt{quadruplets} \\ \hline \\
-     \includegraphics[scale=0.15]{Kate_Winslet_0001.jpg} \includegraphics[scale=0.15]{Nicole_Kidman_0001.jpg} \includegraphics[scale=0.15]{Daniel_Radcliffe_0001.jpg} \includegraphics[scale=0.15]{Kirk_Douglas_0001.jpg} \\
-          \includegraphics[scale=0.15]{Leonardo_DiCaprio_0001.jpg} \includegraphics[scale=0.15]{Angelina_Jolie_0001.jpg} \includegraphics[scale=0.15]{Daryl_Sabara_0001.jpg} \includegraphics[scale=0.15]{Carl_Reiner_0002.jpg} \\
-\end{tabular}
-
-     \caption{}
+\begin{subfigure}
+     \includegraphics[scale=0.2]{quadruplets.pdf}
      \label{fig:quadruplets}
   \end{subfigure}
-  \caption{Different types of supervision.
-\textbf{\ref{fig:full} Full supervision:} we want samples from the same label to be closer than samples from different labels.
-\textbf{\ref{fig:pairs} Pairs supervision:} we want similar samples (1) (the same person) to be closer than dissimilar samples (-1) (different persons).
-\textbf{\ref{fig:quadruplets} Quadruplets supervision:} no labels.
-We want the two leftmost samples (not much age difference) to be closer than the two rightmost samples (bigger age difference).}
+  \caption{Different types of supervision for metric learning.
+(\subref{fig:full}) Full supervision: make samples from the same label (here, identity) closer together than samples from different labels.
+(\subref{fig:pairs}) Pair supervision: make similar pairs (here, corresponding to the same person) to be closer than dissimilar pairs (different persons).
+(\subref{fig:quadruplets}) Quadruplet supervision: make the two leftmost samples (here, people of similar age) to be closer than the two rightmost samples (of bigger age difference).
+(images are the funneled Labeled Faces in the Wild dataset \url{http://vis-www.cs.umass.edu/lfw/} \cite{Huang12}) (\aurelien{designing the 3 figures in inkscape to avoid the ugly tables would be better}}\aurelien{should also acknowledge dataset from which these images are taken}
   \label{fig:contour}
@@
 
+\section{Overview of the Package}
+
+\aurelien{update links, mention scikit-learn-contrib}
+
+\texttt{metric-learn} only relies on core libraries from the SciPy ecosystem: \texttt{numpy}, \texttt{scipy}, and \texttt{scikit-learn}.
+%\footnote{An optional dependency with \texttt{skggm} (\cite{skggm}) can be installed in order to use a more stable solver when using the LSML algorithm from the package.}
+The current release (v.0.5.0) can be installed from the Python Package Index (PyPI), for python 2.7 and 3.5 or later.
+The source code is available on GitHub at \url{http://github.com/scikit-learn-contrib/metric-learn}.
+The development is collaborative and open to all contributors through the usual GitHub framework (issues and pull requests).
+The code is free of use, provided under the MIT licence.
+The interest of the package from the community is clear, since the project has been starred more than 740 times and forked more than 170 times at the time of writing.
+A detailed documentation (including installation guidelines, the description of the algorithms and the API, as well as examples) is available at \url{http://contrib.scikit-learn.org/metric-learn}.
+The quality of the code is ensured by a thorough test coverage (96\% as of July 2019).
+Every new contribution to the code is automatically checked by a continuous integration platform that ensures a minimal test coverage of the added code, a minimal variation in test coverage of the overall project, and syntax formatting with \texttt{flake8}.
+%, which allows easy contribution for incoming developers, which can have their code automatically checked.
+Currently, the following algorithms are implemented in \texttt{metric-learn}: for supervised metric learning: Neighborhood Components Analysis (NCA) (\cite{Goldberger04}), Large Margin Nearest Neighbors (LMNN) (\cite{Weinberger09}), Relative Components Analysis (RCA) (\cite{Shental02}), Local Fisher Discriminant Analysis (LFDA) (\cite{Sugiyama07}) and Metric Learning for Kernel Regression (MLKR) (\cite{Weinberger07}).
+Note that the latter is designed for regression problems and takes as input continuous labels.
+\footnote{\texttt{metric-learn} also provides an algorithm for learning on chunklets of data (groups of points from the same class)}\aurelien{maybe add footnote for RCA?}For pair metric learners: Mahalanobis Metric for Clustering (MMC) (\cite{Xiang08}), Sparse High-Dimensional Metric Learning (SDML) (\cite{Qi09}), Information Theoretic Metric Learning (ITML) (\cite{Davis07}).
+For learning on quadruplets: Metric Learning from Relative Comparisons by Minimizing Squared Residual (LSML) (\cite{Liu12}).
+The details description of these algorithms can be found in the sphinx documentation.
+
+\aurelien{mention flake8 compliance and checks}
+
 \section{Software Architecture and API}
 
-\texttt{metric-learn}'s goal is to provide a unified interface to all metric-learning algorithms, including the ones that have an input format that differs from classical machine learning ones, and allow to combine them with scikit-learn objects.
-As such, all metric-learners from \texttt{metric-learn} inherit from an abstract \texttt{BaseMetricLearner} object.
+\aurelien{mention optimization somewhere?}
+\texttt{metric-learn} provides a unified interface to all metric learning algorithms (including the weakly supervised ones that have an input format that differs from the classic machine learning settings) and is designed to be fully compatible with the functionalities of \texttt{scikit-learn} \citep{scikit-learn}.
+All metric learners in \texttt{metric-learn} inherit from an abstract \texttt{BaseMetricLearner} object, which itself inherits from \texttt{BaseEstimator}.
+All classes inheriting from \texttt{BaseMetricLearner} should implement two methods, \texttt{get\_metric} and \texttt{score\_pairs}: the former returns a function that computes the distance between two points $u$ and $v$.
+The latter method computes the distances between a dataset of pairs of points.
+\aurelien{score\_pairs and get\_metric should be mentioned here (all metric learners should implement these).
+In Mahalanobis, just mention they are implemented accordingly (do not delete: TODO: update accordingly mahalanobis}
 % \todo[inline]{Should I talk about the classes etc ? Not useful since they are not visible by the user right ?}
@@
 
-\texttt{metric-learn}'s API is designed to be fully compatible with \texttt{scikit-learn} API (\cite{scikit-learn}).
+%\texttt{metric-learn}'s API is designed to be fully compatible with \texttt{scikit-learn} API (\cite{scikit-learn}).
 
@@
 
-Supervised metric learners inherit from scikit-learn base class \texttt{TransformerMixin}, just like \texttt{scikit-learn}'s \texttt{LinearDiscriminantAnalysis} for instance, and they behave as such.
+Supervised metric learners inherit from scikit-learn base class \texttt{TransformerMixin}, just like \texttt{scikit-learn}'s \texttt{LinearDiscriminantAnalysis} for instance.
+As such, they are compatible with pipelining (\texttt{sklearn.pipeline.Pipeline} and \texttt{sklearn.pipeline.make\_pipeline}.
 % (they pass all scikit-learn's "estimators checks"). 
 %  They can easily be pipelined with nearest neighbors estimators to improve their performance: \texttt{better\_knn = sklearn.pipeline.make\_pipeline(NCA(), KNeighborsClassifier())} (TODO: sth to explain NCA is in metric-learn)
+\aurelien{mention pipelining}
 
@@
 Tuples can be pairs or quadruplets, depending on the specific algorithm used.
-Below is a code example for pairs supervision:
+(see section~\ref{metriclearning}).
+Say we work with $n_t$ tuples of points with $p$ features each.\\
+\textit{(i) Pairs learners} take as input an array-like \texttt{pairs} of shape $(n_t, 2, p)$, as well as an array-like \texttt{y\_pairs} of shape $(n_t,)$, which for each pair tells whether it contains similar or dissimilar samples.
+In order to \texttt{predict} on incoming pairs whether they are similar or not, they also need to fit a threshold, i.e.
+a reference distance with respect to which the distance between the incoming points will be compared.
+If this distance is higher than the threshold, the pair will be predicted as dissimilar, and if smaller as similar.
+This threshold is automatically fitted at fit time using a heuristic, but can be set manually using the method \texttt{set\_threshold}, or calibrated more specifically on a validation set, in order to optimize any classification score, using the method \texttt{calibrate\_threshold}.\\
+\textit{(ii) Quadruplets learners} work on arrays-like of shape $(n_t, 4, p)$, and no labels array, where for each quadruplet the two first elements are the ones we want to be closer than the two last ones.
+They can also \texttt{predict} whether an incoming quadruplet is in the right order or not, i.e.
+whether the two first elements are indeed more similar that the two last, by comparing the two pairwise distances.
+\aurelien{a bit more details would be welcome (explain both case of pairs and quadruplets). for pairs, need to mention threshold and calibration.}
 
-\begin{minted}{python}
-import numpy as np
-from metric_learn import MMC
-pairs = np.array([[[1.2, 3.2], [2.3, 5.5]], [[4.5, 2.3], [2.1, 2.3]]])
-y_pairs = np.array([1, -1])
-mmc = MMC()
-mmc.fit(pairs, y_pairs)
+\begin{minted}[%linenos=TRUE,
+fontsize=\footnotesize]{python}
+>>> from sklearn.model_selection import train_test_split, cross_val_score
+>>> from sklearn.datasets import fetch_lfw_pairs
+>>> from metric_learn import ITML  
+>>> dataset = fetch_lfw_pairs(resize=0.2)  
+>>> pairs, y_pairs = ((dataset.pairs.reshape(2200, 2, -1) - 127.5) / 255, 
+                      2 * dataset.target - 1)
+>>> pairs, _, y_pairs, _ = train_test_split(pairs, y_pairs, train_size=1000, 
+                                            stratify=y_pairs, random_state=42) 
+>>> itml = ITML(prior='random', random_state=42)  
+>>> cross_val_score(itml, pairs, y_pairs, scoring='accuracy')
+array([0.62275449, 0.65165165, 0.63963964])
 \end{minted}
@@
 
+\aurelien{the snippet does not work, cannot do 3-fold CV on 2 samples..
+maybe we can load a subset of pairs from LFW?
+ideally we should also include the imports so the snippet is self-sufficient.
+to save space, maybe we can split this over 2 columns, with line numbering or something}
+\william{new snippet takes large horizontal space, still split ?}
 This allows to use out of the box \texttt{scikit-learn}'s scoring functions, and therefore all routines for model selection, including the \texttt{GridSearchCV} object.
@@
 
-In conclusion, we presented a scikit-learn compatible package for metric-learning.
-This package is under active development and many improvements are planned to be added to the package.
-First, the scalability of the package could be improved by supporting stochastic solvers (like stochastic gradient descent), and/or to form pairs on the fly and do computations by batch, in order not to load every sample in memory at once.
-We could also take advantage of possible sparse input format to reduce computations.
-More algorithms could also be added to the package, like algorithms learning on triplets, algorithms that can deal with high dimensional data.
-We could also allow metric learn to be used when having multiple labels for each sample, and allow different kind of supervisions, like semi-supervised learning for instance.
-We could also improve the ease of use by adding more examples and adding toy datasets to play with quickly and see the interest of the weaker forms of supervision.
+\texttt{metric-learn} is under active development --- we list here some promising directions to further improve the package.
+Regarding scalability to large datasets, we would like to implement stochastic solvers (stochastic gradient descent and its variants), forming batches of tuples on the fly to avoid loading all data in memory at once.
+We would also like to offer better support to inputs in sparse format.
+Finally, we are looking to incorporate additional algorithms that provide added value to the package, in particular triplet metric learners and algorithms that can deal with multi-label problems \citep{liu15} and high-dimensional data \citep{Liu19}.
+%We could also allow metric learn to be used when having multiple labels for each sample, and allow different kind of supervisions, like semi-supervised learning for instance.
+%We could also improve the ease of use by adding more examples and adding toy datasets to play with quickly and see the interest of the weaker forms of supervision.
 About the overall quality of the software, we intend to improve the continuous integration routines, for instance by ensuring a PEP8 syntax checker is run on every pull request, a continuous integration for the documentation.
@@
 
-We would like to be particularly thankful to Inria for funding two years of development of the package.
-Also we would like to thanks \texttt{scikit-learn} developers for fruitful discussions, assistance for developing the new API, and funding for conference attending.
-\todo[inline]{reformulate + other acknowledgments}
+We are particularly thankful to Inria for funding two years of development of the package.
+We also thank \texttt{scikit-learn} developers from the Inria Parietal team (in particular Gaël Varoquaux, Alexandre Gramfort and Olivier Grisel) for fruitful discussions on the design of the API and for providing some conference funding, as well as scikit-learn-contrib reviewers who integrated the package in the organisation.
+\aurelien{potentially mention contrib reviewers if we are accepted.: include rth and adrinjalali ?}
 

### 2019-08-07 metric-learn-jmlr.tex

@@
 See below for more concrete examples of algorithms learning on tuples.
-\item \textbf{Pair learners.} 
-Another way is to have pair of points as inputs, and a corresponding label indicating whether the two points are similar or not.
-We can then learn a metric that brings points from a similar pair closer together, and points from a dissimilar pair further away from each other.
-In many applications, this can be a much easier type of supervision to collect, in particular when there are many labels.
+\item \textbf{Pair learners} take as input a set of pairs of points with a pair label indicating whether the two points are similar or not.
+They aim to learn a metric that brings points from a similar pair closer together, and points from a dissimilar pair further away from each other.
+Such weak suIn many applications, this weaker supervision can be can be an easier type of supervision to collect, in particular when there are many labels.
 %, since annotators will only need to say whether images are similar or not, and will not have to assign a label in a huge list of those.
-For instance, it is much easier for an annotator to decide whether two images correspond to the same person than to associate an individual image to the right identity among a large number of people (see Figure~\ref{fig:pairs}).
+For instance, it is much easier for an annotator to decide whether two face images correspond to the same person than to match a face to its identity among a large number of people (see Figure~\ref{fig:pairs}).
 % \begin{figure}[H]
@@
 \texttt{metric-learn} provides a unified interface to all metric learning algorithms (including the weakly supervised ones that have an input format that differs from the classic machine learning settings) and is designed to be fully compatible with the functionalities of \texttt{scikit-learn} \citep{scikit-learn}.
-All metric learners in \texttt{metric-learn} inherit from an abstract \texttt{BaseMetricLearner} object, which itself inherits from \texttt{BaseEstimator}.
+All metric learners in \texttt{metric-learn} inherit from an abstract \texttt{BaseMetricLearner} object, which itself inherits from scikit-learn's \texttt{BaseEstimator}.
 All classes inheriting from \texttt{BaseMetricLearner} should implement two methods, \texttt{get\_metric} and \texttt{score\_pairs}: the former returns a function that computes the distance between two points $u$ and $v$.
@@
 
+\william{cf. Nathalie's comment: re-introduce weakly supervised metric laerners}
 \paragraph{Weakly Supervised metric learners} 

### 2019-08-07 metric-learn-jmlr.tex

@@
 They aim to learn a metric that brings points from a similar pair closer together, and points from a dissimilar pair further away from each other.
-Such weak suIn many applications, this weaker supervision can be can be an easier type of supervision to collect, in particular when there are many labels.
+Such weak supervision is often simpler to collect than class labels in applications when there are many labels.
 %, since annotators will only need to say whether images are similar or not, and will not have to assign a label in a huge list of those.
-For instance, it is much easier for an annotator to decide whether two face images correspond to the same person than to match a face to its identity among a large number of people (see Figure~\ref{fig:pairs}).
+For instance, an annotator can easily decide whether two face images correspond to the same person while it can be difficult to match a face to its identity among a large number of people (see Figure~\ref{fig:pairs}).
 % \begin{figure}[H]
@@
 % \end{figure}
-\item \textbf{Quadruplet Learners.}
-The last possibility to learn a metric in \texttt{metric-learn} is to take as input quadruplets of points.
-We will learn a metric that brings the two first points of each quadruplet closer than the two last points.
-This can be used for instance to learn a metric space in which closer points have closer values of a given attributes.
-This is very useful if the attribute is not binary like in pair supervision (e.g.
-same person/not same person), but is rather continuous and difficult to annotate precisely (e.g., the age of a person on an image, see Figure \ref{fig:quadruplets}).
+\item \textbf{Quadruplet Learners} take as input quadruplets of points and aim to learn a metric that brings the two first points of each quadruplet closer than the two last points.
+This can be useful to learn a metric space in which closer points have closer values of an attribute that is not binary like in pair supervision (e.g.
+same person/not same person) but rather continuous and difficult to annotate precisely (e.g., the age of a person on an image, see Figure \ref{fig:quadruplets}).
 %for instance, we want to learn a metric that puts closer together people from approximately the same age.
@@
 
-\begin{figure}[H]
+\begin{figure}[t]
   \begin{subfigure}
@@
 (\subref{fig:quadruplets}) Quadruplet supervision: make the two leftmost samples (here, people of similar age) to be closer than the two rightmost samples (of bigger age difference).
-(images are the funneled Labeled Faces in the Wild dataset \url{http://vis-www.cs.umass.edu/lfw/} \cite{Huang12}) (\aurelien{designing the 3 figures in inkscape to avoid the ugly tables would be better}}\aurelien{should also acknowledge dataset from which these images are taken}
+Images are taken from the Labeled Faces in the Wild dataset \citep{Huang12}.}
   \label{fig:contour}
@@
 
-\aurelien{update links, mention scikit-learn-contrib}
+The current release of \texttt{metric-learn} (v.0.5.0) can be installed from the Python Package Index (PyPI), for python 2.7 and 3.5 or later.
+The source code is available on GitHub at \url{http://github.com/scikit-learn-contrib/metric-learn} and is free of use, provided under the MIT licence.
+A detailed documentation (including installation guidelines, the description of the algorithms and the API, as well as examples) is available at \url{http://contrib.scikit-learn.org/metric-learn}.
 
 \texttt{metric-learn} only relies on core libraries from the SciPy ecosystem: \texttt{numpy}, \texttt{scipy}, and \texttt{scikit-learn}.
-%\footnote{An optional dependency with \texttt{skggm} (\cite{skggm}) can be installed in order to use a more stable solver when using the LSML algorithm from the package.}
-The current release (v.0.5.0) can be installed from the Python Package Index (PyPI), for python 2.7 and 3.5 or later.
-The source code is available on GitHub at \url{http://github.com/scikit-learn-contrib/metric-learn}.
 The development is collaborative and open to all contributors through the usual GitHub framework (issues and pull requests).
-The code is free of use, provided under the MIT licence.
-The interest of the package from the community is clear, since the project has been starred more than 740 times and forked more than 170 times at the time of writing.
-A detailed documentation (including installation guidelines, the description of the algorithms and the API, as well as examples) is available at \url{http://contrib.scikit-learn.org/metric-learn}.
+The interest of the package from the community is clear: it has been included in \texttt{scikit-learn-contrib} and, at the time of writing, has been starred more than 740 times and forked more than 170 times.
 The quality of the code is ensured by a thorough test coverage (96\% as of July 2019).
@@
 %, which allows easy contribution for incoming developers, which can have their code automatically checked.
-Currently, the following algorithms are implemented in \texttt{metric-learn}: for supervised metric learning: Neighborhood Components Analysis (NCA) (\cite{Goldberger04}), Large Margin Nearest Neighbors (LMNN) (\cite{Weinberger09}), Relative Components Analysis (RCA) (\cite{Shental02}), Local Fisher Discriminant Analysis (LFDA) (\cite{Sugiyama07}) and Metric Learning for Kernel Regression (MLKR) (\cite{Weinberger07}).
+
+Currently, \texttt{metric-learn} implements 9 popular metric learning algorithms from the literature.
+Supervised learners include Neighborhood Components Analysis (NCA) (\cite{Goldberger04}), Large Margin Nearest Neighbors (LMNN) (\cite{Weinberger09}), Relative Components Analysis (RCA)\footnote{Strictly speaking, RCA takes as input slightly weaker supervision in the form of \emph{chunklets} (groups of points from the same class).} (\cite{Shental02}), Local Fisher Discriminant Analysis (LFDA) (\cite{Sugiyama07}) and Metric Learning for Kernel Regression (MLKR) (\cite{Weinberger07}).
 Note that the latter is designed for regression problems and takes as input continuous labels.
-\footnote{\texttt{metric-learn} also provides an algorithm for learning on chunklets of data (groups of points from the same class)}\aurelien{maybe add footnote for RCA?}For pair metric learners: Mahalanobis Metric for Clustering (MMC) (\cite{Xiang08}), Sparse High-Dimensional Metric Learning (SDML) (\cite{Qi09}), Information Theoretic Metric Learning (ITML) (\cite{Davis07}).
+Pair learners i Mahalanobis Metric for Clustering (MMC) (\cite{Xiang08}), Sparse High-Dimensional Metric Learning (SDML) (\cite{Qi09}), Information Theoretic Metric Learning (ITML) (\cite{Davis07}).
 For learning on quadruplets: Metric Learning from Relative Comparisons by Minimizing Squared Residual (LSML) (\cite{Liu12}).

### 2019-08-07 metric-learn-jmlr.tex

@@
 Currently, \texttt{metric-learn} implements 9 popular metric learning algorithms from the literature.
-Supervised learners include Neighborhood Components Analysis (NCA) (\cite{Goldberger04}), Large Margin Nearest Neighbors (LMNN) (\cite{Weinberger09}), Relative Components Analysis (RCA)\footnote{Strictly speaking, RCA takes as input slightly weaker supervision in the form of \emph{chunklets} (groups of points from the same class).} (\cite{Shental02}), Local Fisher Discriminant Analysis (LFDA) (\cite{Sugiyama07}) and Metric Learning for Kernel Regression (MLKR) (\cite{Weinberger07}).
+Supervised learners include Neighborhood Components Analysis \citep[NCA,][]{Goldberger04}, Large Margin Nearest Neighbors \citep[LMNN,][]{Weinberger09}, Relative Components Analysis \citep[RCA,][]{Shental02},\footnote{Strictly speaking, RCA takes as input slightly weaker supervision in the form of \emph{chunklets} (groups of points from the same class).} Local Fisher Discriminant Analysis \citep[LFDA,][]{Sugiyama07} and Metric Learning for Kernel Regression \citep[MLKR,][]{Weinberger07}.
 Note that the latter is designed for regression problems and takes as input continuous labels.
-Pair learners i Mahalanobis Metric for Clustering (MMC) (\cite{Xiang08}), Sparse High-Dimensional Metric Learning (SDML) (\cite{Qi09}), Information Theoretic Metric Learning (ITML) (\cite{Davis07}).
-For learning on quadruplets: Metric Learning from Relative Comparisons by Minimizing Squared Residual (LSML) (\cite{Liu12}).
-The details description of these algorithms can be found in the sphinx documentation.
-
-\aurelien{mention flake8 compliance and checks}
+Pair learners include Mahalanobis Metric for Clustering (MMC) (\cite{Xing2002a}), Information Theoretic Metric Learning (ITML) (\cite{Davis07}) and Sparse High-Dimensional Metric Learning (SDML) (\cite{Qi09}).
+Finally, the package implements one quadruplet learner: Metric Learning from Relative Comparisons by Minimizing Squared Residual (LSML) (\cite{Liu12}).
+Detailed descriptions of these algorithms can be found in the package documentation.
 

### 2019-08-07 metric-learn-jmlr.tex

@@
   \begin{subfigure}
-     \includegraphics[scale=0.2]{labels.pdf}
+     \includegraphics[scale=0.5]{labels.pdf}
      \label{fig:full}
@@
  \begin{subfigure}
-     \includegraphics[scale=0.2]{pairs.pdf}
+     \includegraphics[scale=0.5]{pairs.pdf}
      \label{fig:pairs}
@@
 \begin{subfigure}
-     \includegraphics[scale=0.2]{quadruplets.pdf}
+     \includegraphics[scale=0.5]{quadruplets.pdf}
      \label{fig:quadruplets}
@@
 Note that the latter is designed for regression problems and takes as input continuous labels.
-Pair learners include Mahalanobis Metric for Clustering (MMC) (\cite{Xing2002a}), Information Theoretic Metric Learning (ITML) (\cite{Davis07}) and Sparse High-Dimensional Metric Learning (SDML) (\cite{Qi09}).
-Finally, the package implements one quadruplet learner: Metric Learning from Relative Comparisons by Minimizing Squared Residual (LSML) (\cite{Liu12}).
+Pair learners include Mahalanobis Metric for Clustering \citep[MMC,][]{Xing2002a}, Information Theoretic Metric Learning \citep[ITML,][]{Davis07} and Sparse High-Dimensional Metric Learning \citep[SDML,][]{Qi09}.
+Finally, the package implements one quadruplet learner: Metric Learning from Relative Comparisons by Minimizing Squared Residual \citep[LSML,][]{Liu12}.
 Detailed descriptions of these algorithms can be found in the package documentation.
@@
 
-\aurelien{mention optimization somewhere?}
-\texttt{metric-learn} provides a unified interface to all metric learning algorithms (including the weakly supervised ones that have an input format that differs from the classic machine learning settings) and is designed to be fully compatible with the functionalities of \texttt{scikit-learn} \citep{scikit-learn}.
+\aurelien{mention optimization somewhere? rule of thumb opt parameters?}
+\texttt{metric-learn} provides a unified interface to all metric learning algorithms, including weakly supervised ones (pair and quadruplet learners) whose input format differs from the classic machine learning setting.
+It is designed to be fully compatible with the functionalities of \texttt{scikit-learn} \citep{scikit-learn}.
+We explain here how package is structured.
 All metric learners in \texttt{metric-learn} inherit from an abstract \texttt{BaseMetricLearner} object, which itself inherits from scikit-learn's \texttt{BaseEstimator}.

### 2019-08-07 metric-learn-jmlr.tex
Updates from Overleaf

@@
 \usepackage{graphicx}
-\usepackage{svg}
 \usepackage{caption}
@@
 %\renewcommand{\aurelien}[1]{}
+%\renewcommand{\william}[1]{}
 
@@
 % , which is one interesting feature allowed by the metric learning paradigm. 
-Its API is compatible with scikit-learn, the leading library for machine learning in Python, allowing for streamlined model selection, evaluation, and pipelining with other estimators.
+Its API is compatible with \texttt{scikit-learn} \citep{scikit-learn}, the leading library for machine learning in Python, allowing for streamlined model selection, evaluation, and pipelining with other estimators.
 The project is developed collaboratively through GitHub, its code is thoroughly tested using continuous integration tools.
@@
 
-\begin{figure}[t]
-  \begin{subfigure}
-     \includegraphics[scale=0.5]{labels.pdf}
-     \label{fig:full}
-\end{subfigure}
- \begin{subfigure}
-     \includegraphics[scale=0.5]{pairs.pdf}
-     \label{fig:pairs}
-  \end{subfigure}
-\begin{subfigure}
-     \includegraphics[scale=0.5]{quadruplets.pdf}
-     \label{fig:quadruplets}
-  \end{subfigure}
-  \caption{Different types of supervision for metric learning.
+% \begin{figure}[h]
+% \centering
+%   \begin{subfigure}[b]\centering
+%      \includegraphics[scale=0.5]{labels.pdf}
+%      \label{fig:full}
+%   \end{subfigure}
+% \begin{subfigure}[b]
+% \centering
+%      \includegraphics[scale=0.5]{pairs.pdf}
+%      \label{fig:pairs}
+%   \end{subfigure}
+% \begin{subfigure}[b]
+% \centering
+%      \includegraphics[scale=0.5]{quadruplets.pdf}
+%      \label{fig:quadruplets}
+%   \end{subfigure}
+%   \caption{Different types of supervision for metric learning.
 (\subref{fig:full}) Full supervision: make samples from the same label (here, identity) closer together than samples from different labels.
@@
 Images are taken from the Labeled Faces in the Wild dataset \citep{Huang12}.}
-  \label{fig:contour}
+%   \label{fig:contour}
+% \end{figure}
+
+% use t to put figure on top (preferred) Thanks ! 
+\begin{figure}[ht]
+    \centering
+    \begin{subfigure}[t]{0.25\textwidth} % "0.45" donne ici la largeur de l'image
+        \centering \includegraphics[scale=0.45]{labels.pdf}
+        \caption{labels supervision}\label{fig:full}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.25\textwidth}
+        \centering \includegraphics[scale=0.45]{pairs.pdf}
+        \caption{pairwise supervision}\label{fig:pairs}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.40\textwidth}
+        \centering \includegraphics[scale=0.45]{quadruplets.pdf}
+        \caption{quadruplet-wise supervision}\label{fig:quadruplets}
+    \end{subfigure}
+    \caption{(Best seen on computer) Different types of supervision for metric learning.
+(\subref{fig:full}) Full supervision: make samples from the same label (here, identity) closer together than samples from different labels.
+(\subref{fig:pairs}) Pair supervision: make similar pairs (here, corresponding to the same person) to be closer than dissimilar pairs (different persons).
+(\subref{fig:quadruplets}) Quadruplet supervision: make the two leftmost samples (here, people of similar age) to be closer than the two rightmost samples (of bigger age difference).
+Images are taken from the Labeled Faces in the Wild dataset \citep{Huang12}.}\label{fig:flowers}
 \end{figure}
@@
 The source code is available on GitHub at \url{http://github.com/scikit-learn-contrib/metric-learn} and is free of use, provided under the MIT licence.
+\texttt{metric-learn} only relies on core libraries from the SciPy ecosystem: \texttt{numpy}, \texttt{scipy}, and \texttt{scikit-learn}.
 A detailed documentation (including installation guidelines, the description of the algorithms and the API, as well as examples) is available at \url{http://contrib.scikit-learn.org/metric-learn}.
 
-\texttt{metric-learn} only relies on core libraries from the SciPy ecosystem: \texttt{numpy}, \texttt{scipy}, and \texttt{scikit-learn}.
-The development is collaborative and open to all contributors through the usual GitHub framework (issues and pull requests).
+The development is collaborative and open to all contributors through the usual GitHub workflow (issues and pull requests).
 The interest of the package from the community is clear: it has been included in \texttt{scikit-learn-contrib} and, at the time of writing, has been starred more than 740 times and forked more than 170 times.
@@
 %, which allows easy contribution for incoming developers, which can have their code automatically checked.
+\william{a bit repetitive with introduction?}
 
@@
 \aurelien{mention optimization somewhere? rule of thumb opt parameters?}
-\texttt{metric-learn} provides a unified interface to all metric learning algorithms, including weakly supervised ones (pair and quadruplet learners) whose input format differs from the classic machine learning setting.
-It is designed to be fully compatible with the functionalities of \texttt{scikit-learn} \citep{scikit-learn}.
-We explain here how package is structured.
-All metric learners in \texttt{metric-learn} inherit from an abstract \texttt{BaseMetricLearner} object, which itself inherits from scikit-learn's \texttt{BaseEstimator}.
-All classes inheriting from \texttt{BaseMetricLearner} should implement two methods, \texttt{get\_metric} and \texttt{score\_pairs}: the former returns a function that computes the distance between two points $u$ and $v$.
-The latter method computes the distances between a dataset of pairs of points.
-\aurelien{score\_pairs and get\_metric should be mentioned here (all metric learners should implement these).
-In Mahalanobis, just mention they are implemented accordingly (do not delete: TODO: update accordingly mahalanobis}
+\texttt{metric-learn} provides a unified interface to all metric learning algorithms, including 
+%weakly supervised ones (pair and quadruplet learners)
+those whose input format differs from the classic machine learning setting.
+It is designed to be fully compatible with the functionalities of \texttt{scikit-learn}.
+We briefly explain how the package is structured.
+
+All metric learners inherit from an abstract \texttt{BaseMetricLearner} class, which itself inherits from \texttt{scikit-learn}'s \texttt{BaseEstimator}.
+All classes inheriting from \texttt{BaseMetricLearner} should implement two methods, \texttt{get\_metric} (returning a function that computes the distance, which can be plugged into \texttt{scikit-learn} estimators like \texttt{KMeansClustering}) and \texttt{score\_pairs} (returning the distances between a set of pairs of points in a 3D array format).
 % \todo[inline]{Should I talk about the classes etc ? Not useful since they are not visible by the user right ?}
 % \todo[inline]{I think it can be nice, this is not documentation, but explaining how things are coded}
-
 %\texttt{metric-learn}'s API is designed to be fully compatible with \texttt{scikit-learn} API (\cite{scikit-learn}). 
-
-\paragraph{Mahalanobis metric learning}
-
-Since they learn a Mahalanobis matrix, all algorithm also inherit from the \texttt{MahalanobisMixin} interface.
-This latter can return the mahalanobis matrix with the method \texttt{get\_mahalanobis\_matrix}.
+Mahalanobis distance learning algorithms also inherit from a \texttt{MahalanobisMixin} interface, which has an attribute \texttt{components\_} corresponding to the transformation matrix $L$ of the Mahalanobis distance.
+\texttt{MahalanobisMixin} implements \texttt{get\_metric} and \texttt{score\_pairs} accordingly as well as a few additional methods.
+In particular, \texttt{transform} allows to transform data using \texttt{components\_}, and \texttt{get\_mahalanobis\_matrix} returns the Mahalanobis matrix $M=L^TL$.
+% This latter can return the mahalanobis matrix with the method \texttt{get\_mahalanobis\_matrix}.
 Like scikit-learn's algorithms, it can also return a \texttt{components\_} attributes that is the transformation matrix \texttt{L}.
@@
 
-\paragraph{Supervised metric learners}
-
-Supervised metric learners inherit from scikit-learn base class \texttt{TransformerMixin}, just like \texttt{scikit-learn}'s \texttt{LinearDiscriminantAnalysis} for instance.
-As such, they are compatible with pipelining (\texttt{sklearn.pipeline.Pipeline} and \texttt{sklearn.pipeline.make\_pipeline}.
+Supervised metric learners further inherit from \texttt{scikit-learn}'s base class \texttt{TransformerMixin}, just like \texttt{sklearn.LinearDiscriminantAnalysis} for instance.
+As such, they are compatible with pipelining with other estimators through \texttt{sklearn.pipeline.Pipeline} and \texttt{sklearn.pipeline.make\_pipeline}.
 % (they pass all scikit-learn's "estimators checks"). 
 %  They can easily be pipelined with nearest neighbors estimators to improve their performance: \texttt{better\_knn = sklearn.pipeline.make\_pipeline(NCA(), KNeighborsClassifier())} (TODO: sth to explain NCA is in metric-learn)
-\aurelien{mention pipelining}
+% \aurelien{mention pipelining}
 
-\william{cf. Nathalie's comment: re-introduce weakly supervised metric laerners}
-\paragraph{Weakly Supervised metric learners} 
+% \william{cf. Nathalie's comment: re-introduce weakly supervised metric laerners}
 % Algorithms that learn on pairs or quadruplets take as input 3D array of shape \texttt{(n\_samples, t, n\_features)} (where...).
@@
 % Note that \texttt{scikit-learn} accepts arrays in a broad sense, in what they call "array-like" objects. \texttt{metric-learn} too accepts "tuple-like" objects.
-Weakly supervised learning algorithms are similar to \texttt{scikit-learn} classifiers, but they \texttt{fit} on tuples that can be given as 3D arrays, instead of 2D arrays-like of points \texttt{X}.
+Weakly supervised algorithms learning from pairs or quadruplets are similar to \texttt{scikit-learn} classifiers, but they \texttt{fit} on a set of tuples.
+This is passed a 3D array (rather than a 2D array of points).
 They have a \texttt{predict} method, that gives a prediction for each given tuple (e.g.
-for pairs it predicts whether the two points in a pair or similar or not).
+for pairs it predicts whether the two points in a pair are similar or not).
 Tuples can be pairs or quadruplets, depending on the specific algorithm used.
-(see section~\ref{metriclearning}).
 Say we work with $n_t$ tuples of points with $p$ features each.\\
-\textit{(i) Pairs learners} take as input an array-like \texttt{pairs} of shape $(n_t, 2, p)$, as well as an array-like \texttt{y\_pairs} of shape $(n_t,)$, which for each pair tells whether it contains similar or dissimilar samples.
+\textit{(i) Pairs learners} take as input an array-like \texttt{pairs} of shape $(n_t, 2, p)$, as well as an array-like \texttt{y} of shape $(n_t,)$, which for each pair tells whether it contains similar or dissimilar samples.
 In order to \texttt{predict} on incoming pairs whether they are similar or not, they also need to fit a threshold, i.e.
@@
 fontsize=\footnotesize]{python}
->>> from sklearn.model_selection import train_test_split, cross_val_score
->>> from sklearn.datasets import fetch_lfw_pairs
->>> from metric_learn import ITML  
->>> dataset = fetch_lfw_pairs(resize=0.2)  
->>> pairs, y_pairs = ((dataset.pairs.reshape(2200, 2, -1) - 127.5) / 255, 
-                      2 * dataset.target - 1)
->>> pairs, _, y_pairs, _ = train_test_split(pairs, y_pairs, train_size=1000, 
-                                            stratify=y_pairs, random_state=42) 
->>> itml = ITML(prior='random', random_state=42)  
->>> cross_val_score(itml, pairs, y_pairs, scoring='accuracy')
-array([0.62275449, 0.65165165, 0.63963964])
+>>> import sklearn; import metric_learn; import numpy; numpy.random.seed(42)
+>>> _, pairs, y, *_ = sklearn.datasets.fetch_lfw_pairs(resize=0.2).values()
+>>> pairs, y = sklearn.utils.shuffle((pairs.reshape(2200,2,-1)-127.5)/255, 2*y-1) 
+>>> itml = metric_learn.ITML(prior='random')
+>>> sklearn.model_selection.cross_val_score(itml, pairs, y, scoring='accuracy')
 \end{minted}
 % mmc.predict(pairs)
+\william{warning: the $*\_$ syntax is only compatible in python 3 I think, is it OK ? }
 
-\aurelien{the snippet does not work, cannot do 3-fold CV on 2 samples..
-maybe we can load a subset of pairs from LFW?
-ideally we should also include the imports so the snippet is self-sufficient.
-to save space, maybe we can split this over 2 columns, with line numbering or something}
-\william{new snippet takes large horizontal space, still split ?}
 This allows to use out of the box \texttt{scikit-learn}'s scoring functions, and therefore all routines for model selection, including the \texttt{GridSearchCV} object.
-Also note that for every weakly supervised learner, there is a \texttt{\_Supervised} version (e.g.
+Also note that for every pair/quadruplet learner, there is a \texttt{\_Supervised} version (e.g.
 for \texttt{MMC} there is \texttt{MMC\_Supervised}, that is a supervised version that samples pairs of similar samples from intra-class samples, and dissimilar samples from inter-class samples, and fits \texttt{MMC} on these pairs).
@@
 We are particularly thankful to Inria for funding two years of development of the package.
-We also thank \texttt{scikit-learn} developers from the Inria Parietal team (in particular Gaël Varoquaux, Alexandre Gramfort and Olivier Grisel) for fruitful discussions on the design of the API and for providing some conference funding, as well as scikit-learn-contrib reviewers who integrated the package in the organisation.
-\aurelien{potentially mention contrib reviewers if we are accepted.: include rth and adrinjalali ?}
+We also thank \texttt{scikit-learn} developers from the Inria Parietal team (in particular Gaël Varoquaux, Alexandre Gramfort and Olivier Grisel) for fruitful discussions on the design of the API and for providing some conference funding, as well as \texttt{scikit-learn-contrib} reviewers for their useful feedback.
 

### 2019-08-07 metric-learn-jmlr.tex

@@
 % Note that \texttt{scikit-learn} accepts arrays in a broad sense, in what they call "array-like" objects. \texttt{metric-learn} too accepts "tuple-like" objects.
-Weakly supervised algorithms learning from pairs or quadruplets are similar to \texttt{scikit-learn} classifiers, but they \texttt{fit} on a set of tuples.
-This is passed a 3D array (rather than a 2D array of points).
-They have a \texttt{predict} method, that gives a prediction for each given tuple (e.g.
-for pairs it predicts whether the two points in a pair are similar or not).
-Tuples can be pairs or quadruplets, depending on the specific algorithm used.
-Say we work with $n_t$ tuples of points with $p$ features each.\\
-\textit{(i) Pairs learners} take as input an array-like \texttt{pairs} of shape $(n_t, 2, p)$, as well as an array-like \texttt{y} of shape $(n_t,)$, which for each pair tells whether it contains similar or dissimilar samples.
-In order to \texttt{predict} on incoming pairs whether they are similar or not, they also need to fit a threshold, i.e.
-a reference distance with respect to which the distance between the incoming points will be compared.
-If this distance is higher than the threshold, the pair will be predicted as dissimilar, and if smaller as similar.
-This threshold is automatically fitted at fit time using a heuristic, but can be set manually using the method \texttt{set\_threshold}, or calibrated more specifically on a validation set, in order to optimize any classification score, using the method \texttt{calibrate\_threshold}.\\
-\textit{(ii) Quadruplets learners} work on arrays-like of shape $(n_t, 4, p)$, and no labels array, where for each quadruplet the two first elements are the ones we want to be closer than the two last ones.
-They can also \texttt{predict} whether an incoming quadruplet is in the right order or not, i.e.
-whether the two first elements are indeed more similar that the two last, by comparing the two pairwise distances.
-\aurelien{a bit more details would be welcome (explain both case of pairs and quadruplets). for pairs, need to mention threshold and calibration.}
+Weakly supervised algorithms learning from pairs or quadruplets are similar to \texttt{scikit-learn} classifiers, but they \texttt{fit} and \texttt{predict} on a set of tuples.
+This is passed as a 3D array (rather than a 2D array of points).
+Tuples can be pairs or quadruplets depending on the algorithm.
+Pairs learners take as input an array-like \texttt{pairs} of shape \texttt{(n\_pairs, 2, n\_features)}, as well as an array-like \texttt{y} of shape \texttt{(n\_pairs,)} giving labels (similar or dissimilar) for each pair.
+In order to \texttt{predict} the labels of new pairs, one needs to set a threshold on the distance value.
+This threshold can be set manually, or automatically calibrated (at fit time or afterwards on a validation set) to optimize a given performance score such as accuracy or F1-score using the method \texttt{calibrate\_threshold}.
+Quadruplets learners work on array-like of shape \texttt{(n\_quadruplets, 4, n\_features)}, where for each quadruplet the two first elements are the ones we want to be closer than the two last ones.
+They can naturally \texttt{predict} whether a new quadruplet is in the right order or not by comparing the two pairwise distances.
 

### 2019-08-07 metric-learn-jmlr.tex

@@
 They can naturally \texttt{predict} whether a new quadruplet is in the right order or not by comparing the two pairwise distances.
+These design choices allow to use \texttt{scikit-learn}'s scoring functions out of the box, and therefore all routines for model selection, including the \texttt{GridSearchCV} object.
+To illustrate this, we refer to the following code snippet which computes the cross validation score of ITML on pairs from Labeled Faces in the Wild \citep{Huang12}.
+
+% Also note that for every pair/quadruplet learner, there is a \texttt{\_Supervised} version (e.g.
+for \texttt{MMC} there is \texttt{MMC\_Supervised}, that is a supervised version that samples pairs of similar samples from intra-class samples, and dissimilar samples from inter-class samples, and fits \texttt{MMC} on these pairs).
 
@@
 \william{warning: the $*\_$ syntax is only compatible in python 3 I think, is it OK ? }
-
-This allows to use out of the box \texttt{scikit-learn}'s scoring functions, and therefore all routines for model selection, including the \texttt{GridSearchCV} object.
-Also note that for every pair/quadruplet learner, there is a \texttt{\_Supervised} version (e.g.
-for \texttt{MMC} there is \texttt{MMC\_Supervised}, that is a supervised version that samples pairs of similar samples from intra-class samples, and dissimilar samples from inter-class samples, and fits \texttt{MMC} on these pairs).
 

### 2019-08-08 metric-learn-jmlr.tex
Updates from Overleaf

@@
 Distance metric learning, or simply metric learning \citep{Bellet15,Kulis13}, aims at
-automatically constructing task-specific distance metrics from data, in a machine learning manner.
+automatically constructing task-specific distance metrics from data.
 A key advantage of metric learning is that it can be applied beyond the standard supervised learning setting (data points associated with labels), in situations where only weaker forms of supervision are available (e.g., pairs of points that should be similar/dissimilar).
-The learned distance metric can
-then be used to perform various tasks, for instance to retrieve the elements (images, documents) of a database that are semantically closer to a query element.
-It can also be plugged into other machine learning algorithms: popular use-cases include improving the accuracy of nearest neighbors learning models (for classification, regression, anomaly detection...) or biasing the clusters found by clustering algorithms like K-Means towards the intended semantics.
+The learned distance metric can be used to perform retrieval tasks such as finding elements (images, documents) of a database that are semantically closest to a query element.
+It can also be plugged into other machine learning algorithms, for instance to improve the accuracy of nearest neighbors models (for classification, regression, anomaly detection...) or to bias the clusters found by clustering algorithms towards the intended semantics.
 Finally, metric learning can be used to perform dimensionality reduction.
-These use-cases highlight the importance of integrating metric learning into the rest of the machine learning pipeline and tools.
+These use-cases highlight the importance of integrating metric learning with the rest of the machine learning pipeline and tools.
 % \todo[inline]{quick presentation of the package: implements several algorithms, with sklearn-compatible API for model selection, model evaluation, pipelining with other estimators.
@@
 % , which is one interesting feature allowed by the metric learning paradigm. 
-Its API is compatible with \texttt{scikit-learn} \citep{scikit-learn}, the leading library for machine learning in Python, allowing for streamlined model selection, evaluation, and pipelining with other estimators.
-The project is developed collaboratively through GitHub, its code is thoroughly tested using continuous integration tools.
+Its API is compatible with \texttt{scikit-learn} \citep{scikit-learn}, a prominent machine learning library in Python.
+This allows for streamlined model selection, evaluation, and pipelining with other estimators.
+% The project is developed collaboratively through GitHub, its code is thoroughly tested using continuous integration tools.
 It was recently included to \texttt{scikit-learn-contrib}, which hosts high-quality \texttt{scikit-learn}-compatible projects.\footnote{\url{https://github.com/scikit-learn-contrib/scikit-learn-contrib}}
@@
 embedding space of (potentially reduced) dimension \texttt{n\_components}.
-Note that $D_L$ can also be written as $D_L(x, x') = \sqrt{(x - x')^\top M (x - x')}$, where $M = L^\top L$ is called the Mahalanobis matrix.
+Note that $D_L$ can also be written as $D_L(x, x') = \sqrt{(x - x')^\top M (x - x')}$, where we refer to $M = L^\top L$ as the Mahalanobis matrix.
 
@@
 Metric learning algorithms can be classified according to the form of data supervision that they need to learn the metric.
-\texttt{metric-learn} currently implements algorithms that fall into the following categories:
-\begin{itemize}
-\item \textbf{Supervised learners} learn from a standard labeled dataset and essentially aim to bring closer together points from the same class while spreading further away points from a different class.
-For instance, the class may be the identity of a person in an image (see Figure~\ref{fig:full}).
+\texttt{metric-learn} currently implements algorithms that fall into the following categories.
+\emph{Supervised learners} learn from a standard labeled dataset and essentially aim to bring closer together points from the same class while spreading further away points from a different class.
+For instance, data points could be face images and the class could be the identity of the person (see Figure~\ref{fig:full}).
 % \todo[inline]{ces figures ne sont pas très vendeuses.
@@
 See below for more concrete examples of algorithms learning on tuples.
-\item \textbf{Pair learners} take as input a set of pairs of points with a pair label indicating whether the two points are similar or not.
-They aim to learn a metric that brings points from a similar pair closer together, and points from a dissimilar pair further away from each other.
-Such weak supervision is often simpler to collect than class labels in applications when there are many labels.
+\emph{Pair learners} take as input a set of pairs of points with a pair label indicating whether the two points are similar or not.
+They aim to learn a metric that brings pairs of similar points closer together, and pairs of dissimilar points further away from each other.
+Such supervision is often simpler to collect than class labels in applications when there are many labels.
 %, since annotators will only need to say whether images are similar or not, and will not have to assign a label in a huge list of those.
-For instance, an annotator can easily decide whether two face images correspond to the same person while it can be difficult to match a face to its identity among a large number of people (see Figure~\ref{fig:pairs}).
+For instance, a human annotator can often quickly decide whether two face images correspond to the same person (Figure~\ref{fig:pairs}) while matching a face to its identity among many possible people can be difficult.
 % \begin{figure}[H]
@@
 % \end{figure}
-\item \textbf{Quadruplet Learners} take as input quadruplets of points and aim to learn a metric that brings the two first points of each quadruplet closer than the two last points.
-This can be useful to learn a metric space in which closer points have closer values of an attribute that is not binary like in pair supervision (e.g.
-same person/not same person) but rather continuous and difficult to annotate precisely (e.g., the age of a person on an image, see Figure \ref{fig:quadruplets}).
+Finally, \emph{quadruplet learners} take as input quadruplets of points and aim to learn a metric that brings the two first points of each quadruplet closer than the two last points.
+This can be useful to learn a metric space in which closer points have closer values of an attribute that is not binary like in pair supervision but rather continuous and difficult to annotate precisely (e.g., the age of a person on an image, see Figure \ref{fig:quadruplets}).
 %for instance, we want to learn a metric that puts closer together people from approximately the same age.
-\end{itemize}
 
@@
 % use t to put figure on top (preferred) Thanks ! 
-\begin{figure}[ht]
+\begin{figure}[t]
     \centering
@@
         \centering \includegraphics[scale=0.45]{labels.pdf}
-        \caption{labels supervision}\label{fig:full}
+        \caption{class supervision}\label{fig:full}
     \end{subfigure}
@@
         \centering \includegraphics[scale=0.45]{pairs.pdf}
-        \caption{pairwise supervision}\label{fig:pairs}
+        \caption{pair supervision}\label{fig:pairs}
     \end{subfigure}
@@
         \centering \includegraphics[scale=0.45]{quadruplets.pdf}
-        \caption{quadruplet-wise supervision}\label{fig:quadruplets}
+        \caption{quadruplet supervision}\label{fig:quadruplets}
     \end{subfigure}
-    \caption{(Best seen on computer) Different types of supervision for metric learning.
-(\subref{fig:full}) Full supervision: make samples from the same label (here, identity) closer together than samples from different labels.
+    \caption{Different types of supervision for metric learning.
+(\subref{fig:full}) Class supervision: make samples from the same label (here, identity) closer together than samples from different labels.
 (\subref{fig:pairs}) Pair supervision: make similar pairs (here, corresponding to the same person) to be closer than dissimilar pairs (different persons).
-(\subref{fig:quadruplets}) Quadruplet supervision: make the two leftmost samples (here, people of similar age) to be closer than the two rightmost samples (of bigger age difference).
+(\subref{fig:quadruplets}) Quadruplet supervision: make leftmost pairs (here, people of similar age) closer than rightmost pairs (of bigger age difference).
 Images are taken from the Labeled Faces in the Wild dataset \citep{Huang12}.}\label{fig:flowers}
@@
 
-The current release of \texttt{metric-learn} (v.0.5.0) can be installed from the Python Package Index (PyPI), for python 2.7 and 3.5 or later.
+The current release of \texttt{metric-learn} (v.0.5.0) can be installed from the Python Package Index (PyPI), for Python 2.7 and 3.5 or later.
 The source code is available on GitHub at \url{http://github.com/scikit-learn-contrib/metric-learn} and is free of use, provided under the MIT licence.
@@
 A detailed documentation (including installation guidelines, the description of the algorithms and the API, as well as examples) is available at \url{http://contrib.scikit-learn.org/metric-learn}.
+The development is collaborative and open to all contributors through the usual GitHub workflow (issues and pull requests).
+The interest of the community for the package is shown by its recent inclusion in the \texttt{scikit-learn-contrib} organization which hosts high-quality \texttt{scikit-learn}-compatible projects,\footnote{\url{https://github.com/scikit-learn-contrib/scikit-learn-contrib}} and by over 740 stars and 170 forks on GitHub (at the time of writing).
+The quality of the code is ensured by a thorough test coverage (96\% as of July 2019).
+Every new contribution is automatically checked by a continuous integration platform to enforce sufficient test coverage as well as syntax formatting with \texttt{flake8}.
+%, which allows easy contribution for incoming developers, which can have their code automatically checked.
 
-The development is collaborative and open to all contributors through the usual GitHub workflow (issues and pull requests).
-The interest of the package from the community is clear: it has been included in \texttt{scikit-learn-contrib} and, at the time of writing, has been starred more than 740 times and forked more than 170 times.
-The quality of the code is ensured by a thorough test coverage (96\% as of July 2019).
-Every new contribution to the code is automatically checked by a continuous integration platform that ensures a minimal test coverage of the added code, a minimal variation in test coverage of the overall project, and syntax formatting with \texttt{flake8}.
-%, which allows easy contribution for incoming developers, which can have their code automatically checked.
-\william{a bit repetitive with introduction?}
-
-Currently, \texttt{metric-learn} implements 9 popular metric learning algorithms from the literature.
-Supervised learners include Neighborhood Components Analysis \citep[NCA,][]{Goldberger04}, Large Margin Nearest Neighbors \citep[LMNN,][]{Weinberger09}, Relative Components Analysis \citep[RCA,][]{Shental02},\footnote{Strictly speaking, RCA takes as input slightly weaker supervision in the form of \emph{chunklets} (groups of points from the same class).} Local Fisher Discriminant Analysis \citep[LFDA,][]{Sugiyama07} and Metric Learning for Kernel Regression \citep[MLKR,][]{Weinberger07}.
-Note that the latter is designed for regression problems and takes as input continuous labels.
+Currently, \texttt{metric-learn} implements 9 popular metric learning algorithms.
+Supervised learners include Neighborhood Components Analysis \citep[NCA,][]{Goldberger04}, Large Margin Nearest Neighbors \citep[LMNN,][]{Weinberger09}, Relative Components Analysis \citep[RCA,][]{Shental02},\footnote{RCA takes as input slightly weaker supervision in the form of \emph{chunklets} (groups of points of same class).} Local Fisher Discriminant Analysis \citep[LFDA,][]{Sugiyama07} and Metric Learning for Kernel Regression \citep[MLKR,][]{Weinberger07}.
+The latter is designed for regression problems with continuous labels.
 Pair learners include Mahalanobis Metric for Clustering \citep[MMC,][]{Xing2002a}, Information Theoretic Metric Learning \citep[ITML,][]{Davis07} and Sparse High-Dimensional Metric Learning \citep[SDML,][]{Qi09}.
-Finally, the package implements one quadruplet learner: Metric Learning from Relative Comparisons by Minimizing Squared Residual \citep[LSML,][]{Liu12}.
+The package implements one quadruplet learner: Metric Learning from Relative Comparisons by Minimizing Squared Residual \citep[LSML,][]{Liu12}.
 Detailed descriptions of these algorithms can be found in the package documentation.
@@
 \aurelien{mention optimization somewhere? rule of thumb opt parameters?}
-\texttt{metric-learn} provides a unified interface to all metric learning algorithms, including 
+\texttt{metric-learn} provides a unified interface to all metric learning algorithms. %, including 
 %weakly supervised ones (pair and quadruplet learners)
-those whose input format differs from the classic machine learning setting.
+%those whose input format differs from the classic machine learning setting.
 It is designed to be fully compatible with the functionalities of \texttt{scikit-learn}.
-We briefly explain how the package is structured.
-
+% We briefly explain how the package is structured.
 All metric learners inherit from an abstract \texttt{BaseMetricLearner} class, which itself inherits from \texttt{scikit-learn}'s \texttt{BaseEstimator}.
@@
 
-Supervised metric learners further inherit from \texttt{scikit-learn}'s base class \texttt{TransformerMixin}, just like \texttt{sklearn.LinearDiscriminantAnalysis} for instance.
-As such, they are compatible with pipelining with other estimators through \texttt{sklearn.pipeline.Pipeline} and \texttt{sklearn.pipeline.make\_pipeline}.
+Supervised metric learners inherit from \texttt{scikit-learn}'s base class \texttt{TransformerMixin}, just like \texttt{sklearn.LinearDiscriminantAnalysis} for instance.
+As such, they are compatible with pipelining with other estimators (e.g., via \texttt{sklearn.pipeline.Pipeline}).
 % (they pass all scikit-learn's "estimators checks"). 
@@
 % Note that \texttt{scikit-learn} accepts arrays in a broad sense, in what they call "array-like" objects. \texttt{metric-learn} too accepts "tuple-like" objects.
-Weakly supervised algorithms learning from pairs or quadruplets are similar to \texttt{scikit-learn} classifiers, but they \texttt{fit} and \texttt{predict} on a set of tuples.
-This is passed as a 3D array (rather than a 2D array of points).
-Tuples can be pairs or quadruplets depending on the algorithm.
+Weakly supervised algorithms (pairs and quadruplets learners) \texttt{fit} and \texttt{predict} on a set of tuples passed as a 3D array. Tuples can be pairs or quadruplets depending on the algorithm.
 Pairs learners take as input an array-like \texttt{pairs} of shape \texttt{(n\_pairs, 2, n\_features)}, as well as an array-like \texttt{y} of shape \texttt{(n\_pairs,)} giving labels (similar or dissimilar) for each pair.
@@
 To illustrate this, we refer to the following code snippet which computes the cross validation score of ITML on pairs from Labeled Faces in the Wild \citep{Huang12}.
-
 % Also note that for every pair/quadruplet learner, there is a \texttt{\_Supervised} version (e.g.
 for \texttt{MMC} there is \texttt{MMC\_Supervised}, that is a supervised version that samples pairs of similar samples from intra-class samples, and dissimilar samples from inter-class samples, and fits \texttt{MMC} on these pairs).
-
 \begin{minted}[%linenos=TRUE,
 fontsize=\footnotesize]{python}
->>> import sklearn; import metric_learn; import numpy; numpy.random.seed(42)
->>> _, pairs, y, *_ = sklearn.datasets.fetch_lfw_pairs(resize=0.2).values()
->>> pairs, y = sklearn.utils.shuffle((pairs.reshape(2200,2,-1)-127.5)/255, 2*y-1) 
->>> itml = metric_learn.ITML(prior='random')
->>> sklearn.model_selection.cross_val_score(itml, pairs, y, scoring='accuracy')
+>>> from sklearn.datasets import fetch_lfw_pairs
+>>> from sklearn.model_selection import cross_val_score, train_test_split
+>>> from metric_learn import ITML
+>>> _, pairs, y, _, _ = fetch_lfw_pairs(resize=0.2).values()
+>>> pairs, _, y, _ = train_test_split(pairs.reshape(2200,2,-1), 2*y-1) 
+>>> cross_val_score(ITML(), pairs, y, scoring='roc_auc')
 \end{minted}
 % mmc.predict(pairs)
-\william{warning: the $*\_$ syntax is only compatible in python 3 I think, is it OK ? }
+%\william{warning: the $*\_$ syntax is only compatible in python 3 I think, is it OK ? }
+\aurelien{remove $*\_$, use AUC, maybe remove rescaling, potentially try other weakly sup algo, prior random really useful}
+\aurelien{try to have one import per line if we can make that fit}
 
@@
 
-\texttt{metric-learn} is under active development --- we list here some promising directions to further improve the package.
-Regarding scalability to large datasets, we would like to implement stochastic solvers (stochastic gradient descent and its variants), forming batches of tuples on the fly to avoid loading all data in memory at once.
-We would also like to offer better support to inputs in sparse format.
-Finally, we are looking to incorporate additional algorithms that provide added value to the package, in particular triplet metric learners and algorithms that can deal with multi-label problems \citep{liu15} and high-dimensional data \citep{Liu19}.
+\texttt{metric-learn} is under active development.
+We list here promising directions to further improve the package.
+To scale to large datasets, we would like to implement stochastic solvers (SGD and its variants), forming batches of tuples on the fly to avoid loading all data in memory at once.
+% We would also like to offer better support to inputs in sparse format.
+We also plan to incorporate new algorithms that provide added value to the package, in particular triplet metric learners and algorithms that can deal with multi-label \citep{liu15} and high-dimensional problems \citep{Liu19}.
 %We could also allow metric learn to be used when having multiple labels for each sample, and allow different kind of supervisions, like semi-supervised learning for instance.
@@
 About the overall quality of the software, we intend to improve the continuous integration routines, for instance by ensuring a PEP8 syntax checker is run on every pull request, a continuous integration for the documentation.
+\aurelien{if possible, mention algorithms that learn other metrics (similarity, nonlinear, local..}
 
@@
 
-We are particularly thankful to Inria for funding two years of development of the package.
-We also thank \texttt{scikit-learn} developers from the Inria Parietal team (in particular Gaël Varoquaux, Alexandre Gramfort and Olivier Grisel) for fruitful discussions on the design of the API and for providing some conference funding, as well as \texttt{scikit-learn-contrib} reviewers for their useful feedback.
+We are thankful to Inria for funding 2 years of development.
+We also thank \texttt{scikit-learn} developers from the Inria Parietal team (in particular Gaël Varoquaux, Alexandre Gramfort and Olivier Grisel) for fruitful discussions on the design of the API and funding to attend SciPy 2019, as well as \texttt{scikit-learn-contrib} reviewers for their valuable feedback.
 

### 2019-08-10 metric-learn-jmlr.tex
Add CJ’s affiliation Signed-off-by: terrytangyuan <terrytangyuan@gmail.com>

@@
        \name CJ Carey \email perimosocordiae@gmail.com \\
-       \addr \aurelien{CJ must put in his affiliation}
+       \addr Google Inc., United States
        \AND
@@
 % \todo[inline]{probably need to add a sentence explaining what is meant by weak supervision}
-Distance metric learning, or simply metric learning \citep{Bellet15,Kulis13}, aims at
+Distance metric learning, or simply metric learning \citep{Bellet15}, aims at
 automatically constructing task-specific distance metrics from data.
@@
 It was recently included to \texttt{scikit-learn-contrib}, which hosts high-quality \texttt{scikit-learn}-compatible projects.\footnote{\url{https://github.com/scikit-learn-contrib/scikit-learn-contrib}}
-\aurelien{decide if we include or not a few sentences on positioning wrt other packages}
+%\aurelien{decide if we include or not a few sentences on positioning wrt other packages}
 
@@
 
+% use t to put figure on top (preferred) Thanks ! 
+\begin{figure}[t]
+    \centering
+    \begin{subfigure}[t]{0.25\textwidth} % "0.45" donne ici la largeur de l'image
+        \centering \includegraphics[scale=0.45]{labels.pdf}
+        \caption{class supervision}\label{fig:full}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.25\textwidth}
+        \centering \includegraphics[scale=0.45]{pairs.pdf}
+        \caption{pair supervision}\label{fig:pairs}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.40\textwidth}
+        \centering \includegraphics[scale=0.45]{quadruplets.pdf}
+        \caption{quadruplet supervision}\label{fig:quadruplets}
+    \end{subfigure}
+    \caption{Different types of supervision for metric learning %(\subref{fig:full}) Class supervision: make samples from the same label (here, identity) closer together than samples from different labels.
+(\subref{fig:pairs}) Pair supervision: make similar pairs (here, corresponding to the same person) to be closer than dissimilar pairs (different persons).
+(\subref{fig:quadruplets}) Quadruplet supervision: make leftmost pairs (here, people of similar age) closer than rightmost pairs (of bigger age difference).
+    illustrated on face image data taken from the Labeled Faces in the Wild dataset \citep{Huang12}.}\label{fig:flowers}
+\end{figure}
+
 Metric learning algorithms can be classified according to the form of data supervision that they need to learn the metric.
@@
 Finally, \emph{quadruplet learners} take as input quadruplets of points and aim to learn a metric that brings the two first points of each quadruplet closer than the two last points.
-This can be useful to learn a metric space in which closer points have closer values of an attribute that is not binary like in pair supervision but rather continuous and difficult to annotate precisely (e.g., the age of a person on an image, see Figure \ref{fig:quadruplets}).
+This can be used to learn a metric space in which closer points have closer values of an attribute of interest, which may be continuous and difficult to annotate accurately (e.g., the age of a person on an image, see Figure \ref{fig:quadruplets}).
+Quadruplet supervision is also used in problems with a class hierarchy.
 %for instance, we want to learn a metric that puts closer together people from approximately the same age.
@@
 
-% use t to put figure on top (preferred) Thanks ! 
-\begin{figure}[t]
-    \centering
-    \begin{subfigure}[t]{0.25\textwidth} % "0.45" donne ici la largeur de l'image
-        \centering \includegraphics[scale=0.45]{labels.pdf}
-        \caption{class supervision}\label{fig:full}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering \includegraphics[scale=0.45]{pairs.pdf}
-        \caption{pair supervision}\label{fig:pairs}
-    \end{subfigure}
-    \begin{subfigure}[t]{0.40\textwidth}
-        \centering \includegraphics[scale=0.45]{quadruplets.pdf}
-        \caption{quadruplet supervision}\label{fig:quadruplets}
-    \end{subfigure}
-    \caption{Different types of supervision for metric learning.
-(\subref{fig:full}) Class supervision: make samples from the same label (here, identity) closer together than samples from different labels.
-(\subref{fig:pairs}) Pair supervision: make similar pairs (here, corresponding to the same person) to be closer than dissimilar pairs (different persons).
-(\subref{fig:quadruplets}) Quadruplet supervision: make leftmost pairs (here, people of similar age) closer than rightmost pairs (of bigger age difference).
-Images are taken from the Labeled Faces in the Wild dataset \citep{Huang12}.}\label{fig:flowers}
-\end{figure}
+%\aurelien{Fig caption is redundant with the main text. could reduce it.}
 
@@
 The development is collaborative and open to all contributors through the usual GitHub workflow (issues and pull requests).
-The interest of the community for the package is shown by its recent inclusion in the \texttt{scikit-learn-contrib} organization which hosts high-quality \texttt{scikit-learn}-compatible projects,\footnote{\url{https://github.com/scikit-learn-contrib/scikit-learn-contrib}} and by over 740 stars and 170 forks on GitHub (at the time of writing).
+The interest of the community for the package is demonstrated by its recent inclusion in the \texttt{scikit-learn-contrib} organization which hosts high-quality \texttt{scikit-learn}-compatible projects,\footnote{\url{https://github.com/scikit-learn-contrib/scikit-learn-contrib}} and by its more than 740 stars and 170 forks on GitHub (at the time of writing).
 The quality of the code is ensured by a thorough test coverage (96\% as of July 2019).
@@
 
-\aurelien{mention optimization somewhere? rule of thumb opt parameters?}
+%\aurelien{mention optimization somewhere? rule of thumb opt parameters?}
 \texttt{metric-learn} provides a unified interface to all metric learning algorithms. %, including 
@@
 All metric learners inherit from an abstract \texttt{BaseMetricLearner} class, which itself inherits from \texttt{scikit-learn}'s \texttt{BaseEstimator}.
-All classes inheriting from \texttt{BaseMetricLearner} should implement two methods, \texttt{get\_metric} (returning a function that computes the distance, which can be plugged into \texttt{scikit-learn} estimators like \texttt{KMeansClustering}) and \texttt{score\_pairs} (returning the distances between a set of pairs of points in a 3D array format).
+All classes inheriting from \texttt{BaseMetricLearner} should implement two methods, \texttt{get\_metric} (returning a function that computes the distance, which can be plugged into \texttt{scikit-learn} estimators like \texttt{KMeansClustering}) and \texttt{score\_pairs} (returning the distances between a set of pairs of points passed as a 3D array).
 % \todo[inline]{Should I talk about the classes etc ? Not useful since they are not visible by the user right ?}
@@
 % Note that \texttt{scikit-learn} accepts arrays in a broad sense, in what they call "array-like" objects. \texttt{metric-learn} too accepts "tuple-like" objects.
-Weakly supervised algorithms (pairs and quadruplets learners) \texttt{fit} and \texttt{predict} on a set of tuples passed as a 3D array. Tuples can be pairs or quadruplets depending on the algorithm.
-Pairs learners take as input an array-like \texttt{pairs} of shape \texttt{(n\_pairs, 2, n\_features)}, as well as an array-like \texttt{y} of shape \texttt{(n\_pairs,)} giving labels (similar or dissimilar) for each pair.
+Weakly supervised algorithms (pair and quadruplet learners) \texttt{fit} and \texttt{predict} on a set of tuples passed as a 3D array. Tuples can be pairs or quadruplets depending on the algorithm.
+Pair learners take as input an array-like \texttt{pairs} of shape \texttt{(n\_pairs, 2, n\_features)}, as well as an array-like \texttt{y\_pairs} of shape \texttt{(n\_pairs,)} giving labels (similar or dissimilar) for each pair.
 In order to \texttt{predict} the labels of new pairs, one needs to set a threshold on the distance value.
 This threshold can be set manually, or automatically calibrated (at fit time or afterwards on a validation set) to optimize a given performance score such as accuracy or F1-score using the method \texttt{calibrate\_threshold}.
-Quadruplets learners work on array-like of shape \texttt{(n\_quadruplets, 4, n\_features)}, where for each quadruplet the two first elements are the ones we want to be closer than the two last ones.
+Quadruplet learners work on array-like of shape \texttt{(n\_quadruplets, 4, n\_features)}, where for each quadruplet the two first elements are the ones we want to be closer than the two last ones.
 They can naturally \texttt{predict} whether a new quadruplet is in the right order or not by comparing the two pairwise distances.
 These design choices allow to use \texttt{scikit-learn}'s scoring functions out of the box, and therefore all routines for model selection, including the \texttt{GridSearchCV} object.
-To illustrate this, we refer to the following code snippet which computes the cross validation score of ITML on pairs from Labeled Faces in the Wild \citep{Huang12}.
+To illustrate, the following code snippet computes cross validation scores for ITML (with default parameters) on pairs from Labeled Faces in the Wild \citep{Huang12}.
 % Also note that for every pair/quadruplet learner, there is a \texttt{\_Supervised} version (e.g.
@@
 >>> from sklearn.datasets import fetch_lfw_pairs 
->>> from sklearn.model_selection import cross_val_score, train_test_split
+>>> from sklearn.model_selection import cross_validate, train_test_split 
+>>> from sklearn.decomposition import PCA 
 >>> from metric_learn import ITML 
->>> _, pairs, y, _, _ = fetch_lfw_pairs(resize=0.2).values()
->>> pairs, _, y, _ = train_test_split(pairs.reshape(2200,2,-1), 2*y-1) 
->>> cross_val_score(ITML(), pairs, y, scoring='roc_auc')
+>>> pairs, y_pairs = [fetch_lfw_pairs()[key] for key in ['pairs', 'target']]
+>>> pairs = PCA(n_components=25).fit_transform(pairs.reshape(4400, -1)).reshape(-1, 2, 25) 
+>>> pairs, _, y_pairs, _ = train_test_split(pairs, 2*y_pairs-1)  
+>>> cross_validate(ITML(), pairs, y_pairs, scoring='roc_auc', return_train_score=True)
 \end{minted}
@@
 %\william{warning: the $*\_$ syntax is only compatible in python 3 I think, is it OK ? }
-\aurelien{remove $*\_$, use AUC, maybe remove rescaling, potentially try other weakly sup algo, prior random really useful}
-\aurelien{try to have one import per line if we can make that fit}
+%\aurelien{remove $*\_$, use AUC, maybe remove rescaling, potentially try other weakly sup algo, prior random really useful}
+%\aurelien{try to have one import per line if we can make that fit}
 
@@
 
-\section{Conclusion and Future Work}
+\section{Future Work}
 
 \texttt{metric-learn} is under active development.
-We list here promising directions to further improve the package.
+We list here some promising directions to further improve the package.
 To scale to large datasets, we would like to implement stochastic solvers (SGD and its variants), forming batches of tuples on the fly to avoid loading all data in memory at once.
 % We would also like to offer better support to inputs in sparse format.
-We also plan to incorporate new algorithms that provide added value to the package, in particular triplet metric learners and algorithms that can deal with multi-label \citep{liu15} and high-dimensional problems \citep{Liu19}.
+We also plan to incorporate new algorithms that provide added value to the package, in particular some that learn from triplet supervision \citep{Schultz2003a}, can deal with multi-label \citep{liu15} and high-dimensional problems \citep{Liu19}, or learn other forms of metrics \citep[e.g., nonlinear ones, bilinear similarities, and multiple local metrics, see][]{Bellet15}.
 %We could also allow metric learn to be used when having multiple labels for each sample, and allow different kind of supervisions, like semi-supervised learning for instance.
@@
 About the overall quality of the software, we intend to improve the continuous integration routines, for instance by ensuring a PEP8 syntax checker is run on every pull request, a continuous integration for the documentation.
-\aurelien{if possible, mention algorithms that learn other metrics (similarity, nonlinear, local..}
+%\aurelien{if possible, mention algorithms that learn other metrics (similarity, nonlinear, local..}
 

### 2019-08-13 cover-letter-jmlr.tex
Polishing by CJ

\documentclass[a4paper]{article}

\usepackage{a4wide} 
\usepackage{subfigure} 
\usepackage{psfrag}
\usepackage{amsmath}
\usepackage{url}
\usepackage{natbib}
\usepackage{color}
\usepackage{ulem}
\normalem

\usepackage{parskip}

\usepackage{sectsty}
\sectionfont{\large}
\subsectionfont{\normalsize}

\setlength{\textwidth}{6.1in}

\bibliographystyle{unsrt}

\begin{document}

\begin{center}
{\Large \bf metric-learn: Metric Learning Algorithms in Python\\\vspace{.5cm}Cover Letter for JMLR MLOSS Submission}
\vspace{1cm}
\end{center}

Dear JMLR editors,\\

Our submission ``metric-learn: Metric Learning Algorithms in Python'' is intended for the machine learning open source software section. We give here the required information about the software:
\begin{itemize}
    \item Name: metric-learn
    \item License: MIT License
    \item Project URL: \url{https://github.com/scikit-learn-contrib/metric-learn}
    \item Version to be reviewed: v0.5.0\\
\end{itemize}

We look forward to hearing your feedback on our submission.\\

Best regards,\\

William de Vazelhes, CJ Carey, Yuan Tang, Nathalie Vauquier and Aur\'elien Bellet

\end{document}

### 2019-08-13 metric-learn-jmlr.tex
Polishing by CJ

@@
        \name CJ Carey \email perimosocordiae@gmail.com \\
-       \addr Google Inc., United States
+       \addr Google LLC, United States
        \AND
@@
 number of features describing the data, the associated Mahalanobis distance between two points $x$ and $x'$ is defined as $D_L(x, x') = \sqrt{(Lx-Lx')^\top(Lx-Lx')}$.
-This is equivalent to a Euclidean distance after the linear transformation of the feature space defined by $L$ (taking $L$ to be the identity matrix recovers the standard Euclidean distance).
+This is equivalent to Euclidean distance after linear transformation of the feature space defined by $L$.
+Thus, if $L$ is the identity matrix, standard Euclidean distance is recovered.
 Mahalanobis distance metric learning can thus be seen as learning a new
-embedding space of (potentially reduced) dimension \texttt{n\_components}.
+embedding space, with potentially reduced dimension \texttt{n\_components}.
 Note that $D_L$ can also be written as $D_L(x, x') = \sqrt{(x - x')^\top M (x - x')}$, where we refer to $M = L^\top L$ as the Mahalanobis matrix.
@@
 
-Metric learning algorithms can be classified according to the form of data supervision that they need to learn the metric.
+Metric learning algorithms can be categorized according to the form of data supervision they require to learn a metric.
 \texttt{metric-learn} currently implements algorithms that fall into the following categories.
-\emph{Supervised learners} learn from a standard labeled dataset and essentially aim to bring closer together points from the same class while spreading further away points from a different class.
+\emph{Supervised learners} learn from a dataset with one label per training example, aiming to bring together points from the same class while spreading points from different classes.
 For instance, data points could be face images and the class could be the identity of the person (see Figure~\ref{fig:full}). 
@@
 See below for more concrete examples of algorithms learning on tuples.
-\emph{Pair learners} take as input a set of pairs of points with a pair label indicating whether the two points are similar or not.
-They aim to learn a metric that brings pairs of similar points closer together, and pairs of dissimilar points further away from each other.
+\emph{Pair learners} require a set of pairs of points, with each pair labeled to indicate whether the two points are similar or not.
+These methods aim to learn a metric that brings pairs of similar points closer together and pushes pairs of dissimilar points further away from each other.
 Such supervision is often simpler to collect than class labels in applications when there are many labels.
 %, since annotators will only need to say whether images are similar or not, and will not have to assign a label in a huge list of those.
-For instance, a human annotator can often quickly decide whether two face images correspond to the same person (Figure~\ref{fig:pairs}) while matching a face to its identity among many possible people can be difficult.
+For instance, a human annotator can often quickly decide whether two face images correspond to the same person (Figure~\ref{fig:pairs}) while matching a face to its identity among many possible people may be difficult.
 % \begin{figure}[H]
@@
 % \end{figure}
-Finally, \emph{quadruplet learners} take as input quadruplets of points and aim to learn a metric that brings the two first points of each quadruplet closer than the two last points.
-This can be used to learn a metric space in which closer points have closer values of an attribute of interest, which may be continuous and difficult to annotate accurately (e.g., the age of a person on an image, see Figure \ref{fig:quadruplets}).
+Finally, \emph{quadruplet learners} consider 4-tuples of points and aim to learn a metric that brings the two first points of each quadruplet closer than the two last points.
+This can be used to learn a metric space in which closer points are more similar with respect to an attribute of interest, which may be continuous and difficult to annotate accurately (e.g., the age of a person on an image, see Figure \ref{fig:quadruplets}).
 Quadruplet supervision is also used in problems with a class hierarchy.
@@
 The current release of \texttt{metric-learn} (v.0.5.0) can be installed from the Python Package Index (PyPI), for Python 2.7 and 3.5 or later. 
-The source code is available on GitHub at \url{http://github.com/scikit-learn-contrib/metric-learn} and is free of use, provided under the MIT licence.
-\texttt{metric-learn} only relies on core libraries from the SciPy ecosystem: \texttt{numpy}, \texttt{scipy}, and \texttt{scikit-learn}.
-A detailed documentation (including installation guidelines, the description of the algorithms and the API, as well as examples) is available at \url{http://contrib.scikit-learn.org/metric-learn}.
-The development is collaborative and open to all contributors through the usual GitHub workflow (issues and pull requests).
-The interest of the community for the package is demonstrated by its recent inclusion in the \texttt{scikit-learn-contrib} organization which hosts high-quality \texttt{scikit-learn}-compatible projects,\footnote{\url{https://github.com/scikit-learn-contrib/scikit-learn-contrib}} and by its more than 740 stars and 170 forks on GitHub (at the time of writing).
+The source code is available on GitHub at \url{http://github.com/scikit-learn-contrib/metric-learn} and is free to use, provided under the MIT license. 
+\texttt{metric-learn} depends on core libraries from the SciPy ecosystem: \texttt{numpy}, \texttt{scipy}, and \texttt{scikit-learn}.
+Detailed documentation (including installation guidelines, the description of the algorithms and the API, as well as examples) is available at \url{http://contrib.scikit-learn.org/metric-learn}.
+The development is collaborative and open to all contributors through the usual GitHub workflow of issues and pull requests.
+Community interest for the package has been demonstrated by its recent inclusion in the \texttt{scikit-learn-contrib} organization which hosts high-quality \texttt{scikit-learn}-compatible projects,\footnote{\url{https://github.com/scikit-learn-contrib/scikit-learn-contrib}} and by its more than 740 stars and 170 forks on GitHub at the time of writing.
 The quality of the code is ensured by a thorough test coverage (96\% as of July 2019).
@@
 %those whose input format differs from the classic machine learning setting.
-It is designed to be fully compatible with the functionalities of \texttt{scikit-learn}.
+It is designed to be fully compatible with the functionality of \texttt{scikit-learn}.
 % We briefly explain how the package is structured.
 All metric learners inherit from an abstract \texttt{BaseMetricLearner} class, which itself inherits from \texttt{scikit-learn}'s \texttt{BaseEstimator}.
-All classes inheriting from \texttt{BaseMetricLearner} should implement two methods, \texttt{get\_metric} (returning a function that computes the distance, which can be plugged into \texttt{scikit-learn} estimators like \texttt{KMeansClustering}) and \texttt{score\_pairs} (returning the distances between a set of pairs of points passed as a 3D array).
+All classes inheriting from \texttt{BaseMetricLearner} should implement two methods: \texttt{get\_metric} (returning a function that computes the distance, which can be plugged into \texttt{scikit-learn} estimators like \texttt{KMeansClustering}) and \texttt{score\_pairs} (returning the distances between a set of pairs of points passed as a 3D array).
 % \todo[inline]{Should I talk about the classes etc ? Not useful since they are not visible by the user right ?}
@@
 
-Supervised metric learners inherit from \texttt{scikit-learn}'s base class \texttt{TransformerMixin}, just like \texttt{sklearn.LinearDiscriminantAnalysis} for instance.
-As such, they are compatible with pipelining with other estimators (e.g., via \texttt{sklearn.pipeline.Pipeline}).
+Supervised metric learners inherit from \texttt{scikit-learn}'s base class \texttt{TransformerMixin}, the same base class used by \texttt{sklearn.LinearDiscriminantAnalysis} and others.
+As such, they are compatible for pipelining with other estimators via \texttt{sklearn.pipeline.Pipeline}.
 % (they pass all scikit-learn's "estimators checks"). 
@@
 % Note that \texttt{scikit-learn} accepts arrays in a broad sense, in what they call "array-like" objects. \texttt{metric-learn} too accepts "tuple-like" objects.
-Weakly supervised algorithms (pair and quadruplet learners) \texttt{fit} and \texttt{predict} on a set of tuples passed as a 3D array. Tuples can be pairs or quadruplets depending on the algorithm.
+Weakly supervised algorithms (pair and quadruplet learners) \texttt{fit} and \texttt{predict} on a set of tuples passed as a 3-dimensional array.
+Tuples can be pairs or quadruplets depending on the algorithm.
 Pair learners take as input an array-like \texttt{pairs} of shape \texttt{(n\_pairs, 2, n\_features)}, as well as an array-like \texttt{y\_pairs} of shape \texttt{(n\_pairs,)} giving labels (similar or dissimilar) for each pair.
 In order to \texttt{predict} the labels of new pairs, one needs to set a threshold on the distance value.
-This threshold can be set manually, or automatically calibrated (at fit time or afterwards on a validation set) to optimize a given performance score such as accuracy or F1-score using the method \texttt{calibrate\_threshold}.
+This threshold can be set manually or automatically calibrated (at fit time or afterwards on a validation set) to optimize a given score such as accuracy or F1-score using the method \texttt{calibrate\_threshold}.
 Quadruplet learners work on array-like of shape \texttt{(n\_quadruplets, 4, n\_features)}, where for each quadruplet the two first elements are the ones we want to be closer than the two last ones.
-They can naturally \texttt{predict} whether a new quadruplet is in the right order or not by comparing the two pairwise distances.
-These design choices allow to use \texttt{scikit-learn}'s scoring functions out of the box, and therefore all routines for model selection, including the \texttt{GridSearchCV} object.
+They can naturally \texttt{predict} whether a new quadruplet is in the right order by comparing the two pairwise distances.
+These design choices enable use of \texttt{scikit-learn}'s scoring functions out of the box, as well as the standard routines for model selection, including \texttt{GridSearchCV}.
 To illustrate, the following code snippet computes cross validation scores for ITML (with default parameters) on pairs from Labeled Faces in the Wild \citep{Huang12}.

### 2020-06-20 metric-learn-jmlr.tex
minted option to allow arXiv submission
add picture and gitignore
first modifications

@@
 \usepackage{jmlr2e}
-\usepackage{minted}
+\usepackage[cachedir=.,finalizecache]{minted}
 \usepackage[utf8]{inputenc} % allow utf-8 input
@@
 % \aurelien{we could put the presentation of metric learning and supervision as Sec 2, and put the implemented algorithms in the overview of the package section as Section 3}
-
+Other packages for metric-learning already exist, however, they have a different focus: \textit{pyDML} only contains supervised metric learning algorithms, and \textit{pytorch-metric-learning} is focused on deep metric learning, with the pytorch framework.
+\tod[inline]{Talk about the matlab package}
 \section{Background on Metric Learning} \label{metriclearning}
@@
         \caption{pair supervision}\label{fig:pairs}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.40\textwidth}
+        \centering \includegraphics[scale=0.45]{triplets.pdf}
+	\caption{triplet supervision}\label{fig:triplets}
     \end{subfigure}
@@
 %\aurelien{Fig caption is redundant with the main text. could reduce it.}
-
 \section{Overview of the Package}
@@
 Pair learners include Mahalanobis Metric for Clustering \citep[MMC,][]{Xing2002a}, Information Theoretic Metric Learning \citep[ITML,][]{Davis07} and Sparse High-Dimensional Metric Learning \citep[SDML,][]{Qi09}.
-The package implements one quadruplet learner: Metric Learning from Relative Comparisons by Minimizing Squared Residual \citep[LSML,][]{Liu12}.
+The package implements one quadruplet learner: Metric Learning from Relative Comparisons by Minimizing Squared Residual \citep[LSML,][]{Liu12}, and one triplet learner: Sparse Compositional Metric Learning \citep[SCML,][]{Shi15}.
 Detailed descriptions of these algorithms can be found in the package documentation.

### 2020-06-21 metric-learn-jmlr.tex
Update stars, fix some pb with figure, talk about other packages, talk about triplets

--- 2019-08-13 metric-learn-jmlr.tex
@@
 This allows for streamlined model selection, evaluation, and pipelining with other estimators.
+Other packages for metric-learning already exist, however, they have a different focus: \texttt{DistLearn}\footnote{\url{https://www.cs.cmu.edu/~liuy/distlearn.htm}} and \texttt{pyDML}\footnote{\url{http://github.com/jlsuarezdiaz/pydml}} contain almost only fully supervised or unsupervised metric learning algorithms, and \texttt{pytorch-metric-learning}\footnote{\url{http://github.com/KevinMusgrave/pytorch-metric-learning}} is focused on deep metric learning, using the pytorch framework.
 % The project is developed collaboratively through GitHub, its code is thoroughly tested using continuous integration tools.
@@
     \centering
-    \begin{subfigure}[t]{0.25\textwidth} % "0.45" donne ici la largeur de l'image
-        \centering \includegraphics[scale=0.45]{labels.pdf}
+    \begin{subfigure}[t]{0.20\textwidth} % "0.45" donne ici la largeur de l'image
+        \centering \includegraphics[scale=0.35]{labels.pdf}
         \caption{class supervision}\label{fig:full}
     \end{subfigure}
-    \begin{subfigure}[t]{0.25\textwidth}
-        \centering \includegraphics[scale=0.45]{pairs.pdf}
+    \begin{subfigure}[t]{0.22\textwidth}
+        \centering \includegraphics[scale=0.35]{pairs.pdf}
         \caption{pair supervision}\label{fig:pairs}
     \end{subfigure}
-    \begin{subfigure}[t]{0.40\textwidth}
-        \centering \includegraphics[scale=0.45]{quadruplets.pdf}
+    \begin{subfigure}[t]{0.25\textwidth}
+        \centering \includegraphics[scale=0.35]{triplets.pdf}
+	\caption{triplet supervision}\label{fig:triplets}
+    \end{subfigure}
+    \begin{subfigure}[t]{0.30\textwidth}
+        \centering \includegraphics[scale=0.35]{quadruplets.pdf}
         \caption{quadruplet supervision}\label{fig:quadruplets}
@@
 % \end{figure}
+\emph{Triplet learners} consider 3-tuples of points and learn a metric that brings the anchor point and the positive point closer than the anchor and the negative point.
 Finally, \emph{quadruplet learners} consider 4-tuples of points and aim to learn a metric that brings the two first points of each quadruplet closer than the two last points.
-This can be used to learn a metric space in which closer points are more similar with respect to an attribute of interest, which may be continuous and difficult to annotate accurately (e.g., the age of a person on an image, see Figure \ref{fig:quadruplets}).
+Both triplet learners and quadruplets learners can be used to learn a metric space in which closer points are more similar with respect to an attribute of interest, which may be continuous and difficult to annotate accurately (e.g., the age of a person on an image, see Figure \ref{fig:quadruplets}, or their hair color, see Figure \ref{fig:triplets}).
 Quadruplet supervision is also used in problems with a class hierarchy.
@@
 The development is collaborative and open to all contributors through the usual GitHub workflow of issues and pull requests.
-Community interest for the package has been demonstrated by its recent inclusion in the \texttt{scikit-learn-contrib} organization which hosts high-quality \texttt{scikit-learn}-compatible projects,\footnote{\url{https://github.com/scikit-learn-contrib/scikit-learn-contrib}} and by its more than 740 stars and 170 forks on GitHub at the time of writing.
+Community interest for the package has been demonstrated by its recent inclusion in the \texttt{scikit-learn-contrib} organization which hosts high-quality \texttt{scikit-learn}-compatible projects,\footnote{\url{https://github.com/scikit-learn-contrib/scikit-learn-contrib}} and by its more than 990 stars and 200 forks on GitHub at the time of writing.
 The quality of the code is ensured by a thorough test coverage (96\% as of July 2019).
@@
 The package implements one quadruplet learner: Metric Learning from Relative Comparisons by Minimizing Squared Residual \citep[LSML,][]{Liu12}.
+A triplet learner has been recently merged and will be integrated in the next release: Sparse Compositional Metric Learning \citep[SCML,][]{Shi15}.
 Detailed descriptions of these algorithms can be found in the package documentation.

### 2020-06-21 metric-learn-jmlr.tex
Update with triplets

@@
 % Note that \texttt{scikit-learn} accepts arrays in a broad sense, in what they call "array-like" objects. \texttt{metric-learn} too accepts "tuple-like" objects.
-Weakly supervised algorithms (pair and quadruplet learners) \texttt{fit} and \texttt{predict} on a set of tuples passed as a 3-dimensional array.
-Tuples can be pairs or quadruplets depending on the algorithm.
+Weakly supervised algorithms (pair, triplet and quadruplet learners) \texttt{fit} and \texttt{predict} on a set of tuples passed as a 3-dimensional array.
+Tuples can be pairs triplets, or quadruplets depending on the algorithm.
 Pair learners take as input an array-like \texttt{pairs} of shape \texttt{(n\_pairs, 2, n\_features)}, as well as an array-like \texttt{y\_pairs} of shape \texttt{(n\_pairs,)} giving labels (similar or dissimilar) for each pair.
@@
 This threshold can be set manually or automatically calibrated (at fit time or afterwards on a validation set) to optimize a given score such as accuracy or F1-score using the method \texttt{calibrate\_threshold}.
+Triplet learners work on array-like of shape \texttt{(n\_quadruplets, 3, n\_features)}, where for each triplet the second element is the one we want to be closer to the first than the third one.
 Quadruplet learners work on array-like of shape \texttt{(n\_quadruplets, 4, n\_features)}, where for each quadruplet the two first elements are the ones we want to be closer than the two last ones.
-They can naturally \texttt{predict} whether a new quadruplet is in the right order by comparing the two pairwise distances.
+Both triplet and quadruplet learners can naturally \texttt{predict} whether a new triplet/quadruplet is in the right order by comparing the two pairwise distances.
 These design choices enable use of \texttt{scikit-learn}'s scoring functions out of the box, as well as the standard routines for model selection, including \texttt{GridSearchCV}.

### 2020-06-22 metric-learn-jmlr.tex
Update metric-learn-jmlr.tex
Co-authored-by: Yuan Tang <terrytangyuan@gmail.com>

@@
 This allows for streamlined model selection, evaluation, and pipelining with other estimators.
-Other packages for metric-learning already exist, however, they have a different focus: \texttt{DistLearn}\footnote{\url{https://www.cs.cmu.edu/~liuy/distlearn.htm}} and \texttt{pyDML}\footnote{\url{http://github.com/jlsuarezdiaz/pydml}} contain almost only fully supervised or unsupervised metric learning algorithms, and \texttt{pytorch-metric-learning}\footnote{\url{http://github.com/KevinMusgrave/pytorch-metric-learning}} is focused on deep metric learning, using the pytorch framework.
+Other packages for metric learning already exist, however, they have a different focus: \texttt{DistLearn}\footnote{\url{https://www.cs.cmu.edu/~liuy/distlearn.htm}} and \texttt{pyDML}\footnote{\url{http://github.com/jlsuarezdiaz/pydml}} contain almost only fully supervised or unsupervised metric learning algorithms, and \texttt{pytorch-metric-learning}\footnote{\url{http://github.com/KevinMusgrave/pytorch-metric-learning}} is focused on deep metric learning, using the pytorch framework.
 % The project is developed collaboratively through GitHub, its code is thoroughly tested using continuous integration tools.

### 2020-06-22 metric-learn-jmlr.tex
Some fixes and changing the image

@@
     \centering
-    \begin{subfigure}[t]{0.20\textwidth} % "0.45" donne ici la largeur de l'image
+    \begin{subfigure}[t]{0.12\textwidth} % "0.45" donne ici la largeur de l'image
         \centering \includegraphics[scale=0.35]{labels.pdf}
-        \caption{class supervision}\label{fig:full}
+        \caption{class}\label{fig:full}
     \end{subfigure}
-    \begin{subfigure}[t]{0.22\textwidth}
+    \begin{subfigure}[t]{0.20\textwidth}
         \centering \includegraphics[scale=0.35]{pairs.pdf}
-        \caption{pair supervision}\label{fig:pairs}
+        \caption{pair}\label{fig:pairs}
     \end{subfigure}
-    \begin{subfigure}[t]{0.25\textwidth}
+    \begin{subfigure}[t]{0.31\textwidth}
         \centering \includegraphics[scale=0.35]{triplets.pdf}
-	\caption{triplet supervision}\label{fig:triplets}
+	\caption{triplet}\label{fig:triplets}
     \end{subfigure}
-    \begin{subfigure}[t]{0.30\textwidth}
+    \begin{subfigure}[t]{0.31\textwidth}
         \centering \includegraphics[scale=0.35]{quadruplets.pdf}
-        \caption{quadruplet supervision}\label{fig:quadruplets}
+        \caption{quadruplet}\label{fig:quadruplets}
     \end{subfigure}
@@
 The development is collaborative and open to all contributors through the usual GitHub workflow of issues and pull requests.
-Community interest for the package has been demonstrated by its recent inclusion in the \texttt{scikit-learn-contrib} organization which hosts high-quality \texttt{scikit-learn}-compatible projects,\footnote{\url{https://github.com/scikit-learn-contrib/scikit-learn-contrib}} and by its more than 990 stars and 200 forks on GitHub at the time of writing.
+Community interest for the package has been demonstrated by its recent inclusion in the \texttt{scikit-learn-contrib} organization which hosts high-quality \texttt{scikit-learn}-compatible projects,\footnote{\url{https://github.com/scikit-learn-contrib/scikit-learn-contrib}} and by its 1000 stars and 204 forks on GitHub at the time of writing.
 The quality of the code is ensured by a thorough test coverage (96\% as of July 2019).

### 2020-06-22 metric-learn-jmlr.tex
knitpick
Address final comments

@@
 This allows for streamlined model selection, evaluation, and pipelining with other estimators.
-Other packages for metric learning already exist, however, they have a different focus: \texttt{DistLearn}\footnote{\url{https://www.cs.cmu.edu/~liuy/distlearn.htm}} and \texttt{pyDML}\footnote{\url{http://github.com/jlsuarezdiaz/pydml}} contain almost only fully supervised or unsupervised metric learning algorithms, and \texttt{pytorch-metric-learning}\footnote{\url{http://github.com/KevinMusgrave/pytorch-metric-learning}} is focused on deep metric learning, using the pytorch framework.
+Other packages for metric learning already exist, however, they have a different focus: \texttt{pyDML} \footnote{\url{http://github.com/jlsuarezdiaz/pydml}} contains almost only fully supervised or unsupervised metric learning algorithms, and \texttt{pytorch-metric-learning} \footnote{\url{http://github.com/KevinMusgrave/pytorch-metric-learning}} is focused on deep metric learning, using the \texttt{pytorch} \citep{pytorch} framework.
+There is also a Matlab toolkit, \texttt{DistLearn} \footnote{\url{https://www.cs.cmu.edu/~liuy/distlearn.htm}}, that is a webpage gathering several implementations of metric-learning algorithms, however they are mostly fully supervised or unsupervised, and they lack a common interface.
 % The project is developed collaboratively through GitHub, its code is thoroughly tested using continuous integration tools.
@@
 %\aurelien{decide if we include or not a few sentences on positioning wrt other packages}
-
 %, which makes \texttt{metric-learn} stand out from other packages, like \texttt{PyDML} \footnote{\url{https://github.com/jlsuarezdiaz/pyDML}}, which mostly focuses on supervised metric learning algorithms.
@@
         \centering \includegraphics[scale=0.35]{labels.pdf}
-        \caption{class}\label{fig:full}
+        \caption{classes}\label{fig:full}
     \end{subfigure}
@@
         \centering \includegraphics[scale=0.35]{pairs.pdf}
-        \caption{pair}\label{fig:pairs}
+        \caption{pairs}\label{fig:pairs}
     \end{subfigure}
@@
         \centering \includegraphics[scale=0.35]{triplets.pdf}
-	\caption{triplet}\label{fig:triplets}
+	\caption{triplets}\label{fig:triplets}
     \end{subfigure}
@@
         \centering \includegraphics[scale=0.35]{quadruplets.pdf}
-        \caption{quadruplet}\label{fig:quadruplets}
+        \caption{quadruplets}\label{fig:quadruplets}
     \end{subfigure}

### 2020-07-01 cover-letter-jmlr-revision.tex
Updates from Overleaf

--- 2019-08-13 cover-letter-jmlr.tex
@@
 
-Our submission ``metric-learn: Metric Learning Algorithms in Python'' is intended for the machine learning open source software section. We give here the required information about the software:
-\begin{itemize}
-    \item Name: metric-learn
-    \item License: MIT License
-    \item Project URL: \url{https://github.com/scikit-learn-contrib/metric-learn}
-    \item Version to be reviewed: v0.5.0\\
-\end{itemize}
-
-We look forward to hearing your feedback on our submission.\\
+We have the honor to submit the revised version of the manuscript, entitled "metric-learn: Metric Learning Algorithms in Python" for publication in Journal of Machine Learning Research.
+All the comments made by the referee have been addressed.
 

### 2020-07-01 header-PBP.tex
Updates from Overleaf

%%%%% DOCUMENT SETUP
\usepackage[top = 1.5cm, bottom = 1.5cm, right = 1.5cm, left = 1.5cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage[titles]{tocloft}
\usepackage[explicit]{titlesec}
\usepackage{fancybox}
\usepackage[bottom]{footmisc}
\usepackage{fancyhdr}
\usepackage{etex}
\usepackage{lastpage}
\usepackage{setspace}

%%%%% FONTS
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage[sc]{mathpazo}
\usepackage{dsfont}

%%%%%% SYMBOLS
%\usepackage{latexsym, pifont}

%%%%% MATHS
\usepackage{amssymb, amsfonts, amsmath, euler, amscd, amsthm, array}

%%%%%% TABLES
\usepackage{booktabs, tabularx, float, subfig, multicol, multirow, caption, wrapfig}
\setlength{\aboverulesep}{0pt}
\setlength{\belowrulesep}{0pt}
\captionsetup{format = plain, labelfont = normalfont, textfont = {it, small}, labelsep = endash, justification = centerlast}
\usepackage{color, colortbl}
\usepackage[table]{xcolor}

%%%%%% FIGURES
\usepackage{graphicx, epsfig, tikz, subfig}

%%%%%% HYPERREF
\RequirePackage[colorlinks, citecolor = autumnred, urlcolor = autumnred, linkcolor = darkautumnred]{hyperref}

%%%%% LISTS
%\usepackage{enumerate}
\usepackage{enumitem}

%%%%% MISC
\usepackage{pdfpages}
\usepackage{textcomp}
\usepackage{sidecap}
\usepackage{ifthen}

%%%%% TITLES FORMAT
\renewcommand{\thesection}{\Roman{section}}
\renewcommand{\thesubsection}{\arabic{subsection}}
\titleformat{\section}[hang]{\centering\scshape\LARGE\color{bordeau!50!darkautumnred}}{\huge\thesection\,-\,}{0pt}{#1\\[-1em]\textcolor{bordeau!50!darkautumnred}{\hrule height 1.5pt}}
\titleformat{\subsection}[hang]{\scshape\large}{{\bfseries\color{bordeau!50!darkautumnred}\thesubsection \,-\,}}{5pt}{#1}{}
%\titleformat{\paragraph}[hang]{\scshape}{}{5pt}{#1}{}

%%%%%% TITLES FORMAT
%\renewcommand{\thesection}{\huge \roman{section}\,-\,}
%\renewcommand{\thesubsection}{\arabic{subsection}\,-\,}
%\titleformat{\section}[hang]{\scshape\LARGE\color{bordeau}}{\thesection}{0pt}{#1 \textcolor{bordeau}{\titlerule[1.25pt]}}
%\titleformat{\subsection}[hang]{\bfseries\large}{\thesubsection}{5pt}{ #1}

%%%%% COLORS
\definecolor{bordeau}{rgb}{0.46 , 0.06 , 0.06}
\definecolor{gris}{rgb}{0.60 , 0.60 , 0.60}
\definecolor{grisclair}{RGB}{234,234,234}
\definecolor{autumnred}{RGB}{169,56,38}
\colorlet{darkautumnred}{autumnred!120}

%%%%% NEW COMMANDS : SHORTCUTS

%% LATIN ABBREVIATIONS
\newcommand{\ie}{\textit{i.e.}\hspace{4pt}}
\newcommand{\eg}{\textit{e.g.}\hspace{4pt}}
\newcommand{\cf}{\textit{cf.}\hspace{4pt}}

\newcommand{\red}[1]{\textcolor{red}{#1}}

%% MATH SYMBOLS
\newcommand{\opset}[1]{ \overset{\circ}{#1} } % Interior
\newcommand{\wideopset}[1]{ \overset{\circ}{\overgroup{#1}} } % Interior (wide)
\newcommand{\clset}[1]{ \overline{#1} } % Closure
\newcommand{\ginv}[1]{ #1^{\leftarrow} } % Generalized inverse
\newcommand{\Prob}[2][]{ \mathbb{P}_{#1}\left( #2 \right)} % Probability
\newcommand{\Probhat}[2][]{ \widehat{\mathbb{P}}_{#1}\left( #2 \right)} % Probability estimator
\newcommand{\Esp}[2][]{ \mathbb{E}_{#1}\left( #2 \right) } % Expectation
\newcommand{\Esphat}[2][]{ \widehat{\mathbb{E}}_{#1}\left( #2 \right) } % Expectation estimator
\newcommand{\Var}[2][]{ \mathbb{V}_{#1}\left( #2 \right) } % Variance
\newcommand{\Varhat}[2][]{ \widehat{\mathbb{V}}_{#1}\left( #2 \right) } % Variance estimator
\newcommand{\ascv}[1][n]{ \underset{ {#1} \rightarrow \infty }{\longrightarrow} } % Asymptotic convergence
\newcommand{\ascvvg}[1][n]{ \overset{v}{\underset{ {#1} \rightarrow \infty }{\longrightarrow} }} % Vague asymptotic convergence
\newcommand{\ascvlaw}[1][n]{ \overset{\mathcal{D}}{\underset{ {#1} \rightarrow \infty }{\longrightarrow}} } % Asymptotic convergence in distribution
\newcommand{\ascvprob}[1][n]{ \overset{\mathbb{P}}{\underset{ {#1} \rightarrow \infty }{\longrightarrow}} } % Asymptotic convergence in probability
\newcommand{\iid}{\overset{iid}{\leadsto}} % iid distributed
\newcommand{\vlim}[1][n]{\underset{#1 \rightarrow \infty}{v\hspace{-1.5pt}\lim} \,\, } % Vague limit
\newcommand{\indic}[1]{\mathbb{I}\left\{ #1 \right\}} % Indicator function
\newcommand{\COND}{\bigg\vert} % Conditionning (large separator)

%% MATH FONTS
\def\P{\mathcal P}
\def\S{\mathcal S}
\def\RR{\mathbb R}
\newcommand{\F}{{\cal F}}
\newcommand{\G}{{\cal G}}
\newcommand{\X}{{\cal X}}
\newcommand{\D}{{\cal D}}
\newcommand{\cR}{{\cal R}}
\def\C{{\cal C}}
\def\B{{\cal B}}
\def\A{{\cal A}}
\def\N{{\cal N}}
\def\F{{\cal F}}
\def\S{{\cal S}}
\def\L{{\cal L}}
\def\K{{\cal K}}

%%% DOCUMENT INFORMATIONS
% AUTHORS
\newcommand{\theauthor}{}	
\newcommand{\theshortauthor}{}	% Short version optional
\renewcommand*{\author}[2][]{\renewcommand{\theshortauthor}{#1}\renewcommand{\theauthor}{#2}}
% JOURNAL
\newcommand{\thejournal}{}	
\newcommand{\theshortjournal}{}	% Short version optional
\newcommand{\journal}[2][]{\renewcommand{\theshortjournal}{#1}\renewcommand{\thejournal}{#2}}
% SUBMISSION NUMBER
\newcommand{\thesubnb}{}
\newcommand{\subnb}[1]{\renewcommand{\thesubnb}{#1}}
% DATE
\newcommand{\thedate}{}
\renewcommand{\date}[1]{\renewcommand{\thedate}{#1}}
% PAPER TITLE
\newcommand{\thepaper}{}
\newcommand{\theshortpaper}{}
\newcommand{\paper}[2][]{\renewcommand{\theshortpaper}{#1}\renewcommand{\thepaper}{#2}}

%%% LENGHTS
\setlength{\parskip}{4pt}				%Saut de paragraphe
\setlength{\headheight}{15.2pt}			%Hauteur de l'entête

%%% HEADER & FOOTER
\pagestyle{fancy}
\fancyhead{}						
\fancyhead{\begin{tikzpicture}
[remember picture, overlay] \node[anchor = north west, fill = gray!50, minimum width = \paperwidth, text width = .95\paperwidth, text ragged, inner sep = 5pt] at (current page.north west) {\itshape \ifthenelse{\equal{\theshortauthor}{}}{\theauthor}{\theshortauthor} \hfill \ifthenelse{\equal{\theshortpaper}{}}{\thepaper}{\theshortpaper}}; \end{tikzpicture}}
\fancyfoot{}						
\fancyfoot{\begin{tikzpicture}
[remember picture, overlay] \node[anchor = south west, fill = gray!50, minimum width = \paperwidth, text width = .95\paperwidth, text ragged, inner sep = 5pt] at (current page.south west) {{\itshape \thedate} \hfill \thepage /\pageref{LastPage}}; \end{tikzpicture}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

%%% TITLE LAYOUT
\renewcommand{\maketitle}{
%\shorthandoff{:}\shorthandoff{!}
\begin{tikzpicture}
	\node[minimum width = \paperwidth, minimum height = 0.4\paperheight] at (current page.north west) {};
\end{tikzpicture}
\begin{tikzpicture}[remember picture, overlay]
%
%	\node[fill = gray!50, minimum width = \paperwidth, text width = 0.95\paperwidth, inner xsep = 1em, inner ysep = 1em, anchor = north west] (head) at (current page.north west) {\large \scshape \theschool \hfill \thelevel};
%
	\node[fill = gray, minimum width = \paperwidth, inner ysep = 3em, anchor = north west] (title) at ([yshift = -7em]current page.north west) {\begin{minipage}{.95\paperwidth} \begin{spacing}{2}
\centering {\huge \scshape Point-by-point reply to the referee report on \thejournal \ submission No. \thesubnb}
	\end{spacing}\vspace*{-1em}
	\end{minipage}};
%	
%	\node[anchor = base, minimum width = \paperwidth] (course) at  ($(head.south)!0.5!(title.north)$) {\begin{minipage}{\paperwidth} \centering {\scshape \Huge \thecourse} \ifthenelse{\equal{\thesubtitle}{}}{}{{\\\vspace*{2em} \Large \slshape Cours de \theprof}} \end{minipage}};
%	
%	\node[fill = gray!50, minimum width = \paperwidth, text width = .95\paperwidth, inner xsep = 1em, inner ysep = 1em, anchor = south west] (foot) at (current page.south west) {\large  \scshape \theauthor \hfill \thedate};
%	
	\node[anchor = north west, minimum width = \paperwidth] (auth) at ([xshift = 2em, yshift = -3em]title.south west) {\begin{minipage}{\paperwidth} {\scshape\color{bordeau}\Large Authors:} \hspace{.5em} {\large\textbf{\theauthor}} \end{minipage}};
	\node[anchor = north west, minimum width = \paperwidth] (pap) at ([yshift = -1em]auth.south west) {\begin{minipage}{\paperwidth} {\scshape\color{bordeau}\Large Title:} \hspace{.5em} {\large\textbf{\thepaper}} \end{minipage}};
%
\end{tikzpicture}
%\shorthandon{:}\shorthandon{!}
}

### 2020-07-01 metric-learn-jmlr.tex
Updates from Overleaf

@@
 %\renewcommand{\william}[1]{}
+\usepackage{xspace}
+\newcommand{\MATLAB}{\textsc{Matlab}\xspace}
 
@@
        \name Yuan Tang \email terrytangyuan@gmail.com \\
-       \addr Ant Financial, United States
+       \addr Ant Financial Services Group, United States
        \AND
@@
 This allows for streamlined model selection, evaluation, and pipelining with other estimators.
-Other packages for metric learning already exist, however, they have a different focus: \texttt{pyDML} \footnote{\url{http://github.com/jlsuarezdiaz/pydml}} contains almost only fully supervised or unsupervised metric learning algorithms, and \texttt{pytorch-metric-learning} \footnote{\url{http://github.com/KevinMusgrave/pytorch-metric-learning}} is focused on deep metric learning, using the \texttt{pytorch} \citep{pytorch} framework.
-There is also a Matlab toolkit, \texttt{DistLearn} \footnote{\url{https://www.cs.cmu.edu/~liuy/distlearn.htm}}, that is a webpage gathering several implementations of metric-learning algorithms, however they are mostly fully supervised or unsupervised, and they lack a common interface.
+
+\paragraph{Positioning with respect to other packages.} Many metric learning algorithms were originally implemented by their authors in \MATLAB without a common API convention.\footnote{See \url{https://www.cs.cmu.edu/~liuy/distlearn.htm} for a list of \MATLAB implementations.}  In R, the package \texttt{dml} \citep{Tang18} implements several metric learning algorithms with a unified interface but is not tightly integrated with any general-purpose machine learning library.
+In Python, \texttt{pyDML}\footnote{\url{http://github.com/jlsuarezdiaz/pydml}} contains mainly fully supervised and unsupervised algorithms, while \texttt{pytorch-metric-learning}\footnote{\url{http://github.com/KevinMusgrave/pytorch-metric-learning}} focuses on deep metric learning using the \texttt{pytorch} framework \citep{pytorch}.
+
+% Other packages for metric learning already exist, however, they have a different focus: \texttt{pyDML} \footnote{\url{http://github.com/jlsuarezdiaz/pydml}} contains almost only fully supervised or unsupervised metric learning algorithms, and \texttt{pytorch-metric-learning} \footnote{\url{http://github.com/KevinMusgrave/pytorch-metric-learning}} is focused on deep metric learning, using the \texttt{pytorch} \citep{pytorch} framework.
+There are also resources for metric learning in other languages like Matlab and R.
+In Matlab, the toolkit \texttt{DistLearn} \footnote{\url{https://www.cs.cmu.edu/~liuy/distlearn.htm}}is a webpage gathering several implementations of metric-learning algorithms, however they are mostly fully supervised or unsupervised, and they lack a common interface.
+In R, the package \texttt{dml} \citep{Tang18} implements several metric-learning algorithms with a unified interface.
+
 % The project is developed collaboratively through GitHub, its code is thoroughly tested using continuous integration tools.
@@
 % \end{figure}
-\emph{Triplet learners} consider 3-tuples of points and learn a metric that brings the anchor point and the positive point closer than the anchor and the negative point.
+\emph{Triplet learners} consider 3-tuples of points and learn a metric that brings the first (\emph{anchor}) point of each triplet closer to the second point than to the third one.
 Finally, \emph{quadruplet learners} consider 4-tuples of points and aim to learn a metric that brings the two first points of each quadruplet closer than the two last points.
-Both triplet learners and quadruplets learners can be used to learn a metric space in which closer points are more similar with respect to an attribute of interest, which may be continuous and difficult to annotate accurately (e.g., the age of a person on an image, see Figure \ref{fig:quadruplets}, or their hair color, see Figure \ref{fig:triplets}).
-Quadruplet supervision is also used in problems with a class hierarchy.
+Both triplet and quadruplets learners can be used to learn a metric space where closer points are more similar with respect to an attribute of interest, in particular when this attribute is continuous and/or difficult to annotate accurately (e.g., the hair color of a person on an image, see Figure \ref{fig:triplets}, or the age of a person, see Figure \ref{fig:quadruplets}).
+Triplet and quadruplet supervision can also be used in problems with a class hierarchy.
 %for instance, we want to learn a metric that puts closer together people from approximately the same age.
@@
 
-The current release of \texttt{metric-learn} (v.0.5.0) can be installed from the Python Package Index (PyPI), for Python 2.7 and 3.5 or later. 
+The current release of \texttt{metric-learn} (v0.6.0) can be installed from the Python Package Index (PyPI) and conda-forge, for Python 3.6 or later.\footnote{Support for Python 2.7 and 3.5 was dropped in the last release.}
 The source code is available on GitHub at \url{http://github.com/scikit-learn-contrib/metric-learn} and is free to use, provided under the MIT license. 
@@
 The development is collaborative and open to all contributors through the usual GitHub workflow of issues and pull requests.
-Community interest for the package has been demonstrated by its recent inclusion in the \texttt{scikit-learn-contrib} organization which hosts high-quality \texttt{scikit-learn}-compatible projects,\footnote{\url{https://github.com/scikit-learn-contrib/scikit-learn-contrib}} and by its 1000 stars and 204 forks on GitHub at the time of writing.
-The quality of the code is ensured by a thorough test coverage (96\% as of July 2019).
+Community interest for the package has been demonstrated by its recent inclusion in the \texttt{scikit-learn-contrib} organization which hosts high-quality \texttt{scikit-learn}-compatible projects,\footnote{\url{https://github.com/scikit-learn-contrib/scikit-learn-contrib}} and by its more than 1000 stars and 200 forks on GitHub at the time of writing.
+The quality of the code is ensured by a thorough test coverage (97\% as of June 2020).
 Every new contribution is automatically checked by a continuous integration platform to enforce sufficient test coverage as well as syntax formatting with \texttt{flake8}.
@@
 
-Currently, \texttt{metric-learn} implements 9 popular metric learning algorithms.
+Currently, \texttt{metric-learn} implements 10 popular metric learning algorithms.
 Supervised learners include Neighborhood Components Analysis \citep[NCA,][]{Goldberger04}, Large Margin Nearest Neighbors \citep[LMNN,][]{Weinberger09}, Relative Components Analysis \citep[RCA,][]{Shental02},\footnote{RCA takes as input slightly weaker supervision in the form of \emph{chunklets} (groups of points of same class).} Local Fisher Discriminant Analysis \citep[LFDA,][]{Sugiyama07} and Metric Learning for Kernel Regression \citep[MLKR,][]{Weinberger07}.
@@
 Pair learners include Mahalanobis Metric for Clustering \citep[MMC,][]{Xing2002a}, Information Theoretic Metric Learning \citep[ITML,][]{Davis07} and Sparse High-Dimensional Metric Learning \citep[SDML,][]{Qi09}.
-The package implements one quadruplet learner: Metric Learning from Relative Comparisons by Minimizing Squared Residual \citep[LSML,][]{Liu12}.
-A triplet learner has been recently merged and will be integrated in the next release: Sparse Compositional Metric Learning \citep[SCML,][]{Shi15}.
+Finally, the package implements one triplet learner and one quadruplet learner: Sparse Compositional Metric Learning \citep[SCML,][]{Shi15} and Metric Learning from Relative Comparisons by Minimizing Squared Residual \citep[LSML,][]{Liu12}.
+% A triplet learner has been recently merged and will be integrated in the next release: Sparse Compositional Metric Learning \citep[SCML,][]{Shi15}.
 Detailed descriptions of these algorithms can be found in the package documentation.
@@
 As such, they are compatible for pipelining with other estimators via \texttt{sklearn.pipeline.Pipeline}.
+To illustrate, the following code snippet trains a \texttt{Pipeline} composed of LMNN followed by a k-nearest neighbors classifier on the UCI Wine dataset, with the hyperparameters selected with a grid-search.
+Any other supervised metric learner can be used in place of LMNN.
+% Also note that for every pair/quadruplet learner, there is a \texttt{\_Supervised} version (e.g.
+for \texttt{MMC} there is \texttt{MMC\_Supervised}, that is a supervised version that samples pairs of similar samples from intra-class samples, and dissimilar samples from inter-class samples, and fits \texttt{MMC} on these pairs).
+%\begin{minted}[%linenos=TRUE,
+%fontsize=\footnotesize]{python}
+%>>> from sklearn.datasets import fetch_lfw_pairs 
+%>>> from sklearn.model_selection import cross_validate, train_test_split 
+%>>> from sklearn.decomposition import PCA 
+%>>> from metric_learn import ITML 
+%>>> pairs, y_pairs = [fetch_lfw_pairs()[key] for key in ['pairs', 'target']]
+%>>> pairs = PCA(n_components=25).fit_transform(pairs.reshape(4400, -1)).reshape(-1, 2, 25) 
+%>>> pairs, _, y_pairs, _ = train_test_split(pairs, 2*y_pairs-1)  
+%>>> cross_validate(ITML(), pairs, y_pairs, scoring='roc_auc', return_train_score=True)
+%\end{minted}
+\begin{minted}[%linenos=TRUE,
+fontsize=\footnotesize]{python}
+from sklearn.datasets import load_wine
+from sklearn.neighbors import KNeighborsClassifier
+from sklearn.model_selection import train_test_split, GridSearchCV
+from sklearn.pipeline import Pipeline
+from metric_learn import LMNN
+
+X_train, X_test, y_train, y_test = train_test_split(*load_wine(return_X_y=True))
+lmnn_knn = Pipeline(steps=[('lmnn', LMNN()), ('knn', KNeighborsClassifier())])
+parameters = {'lmnn__k':[1, 2], 'knn__n_neighbors':[1, 2]}
+grid_lmnn_knn = GridSearchCV(lmnn_knn, parameters, cv=3, n_jobs=-1, verbose=True)
+grid_lmnn_knn.fit(X_train, y_train)
+grid_lmnn_knn.score(X_test, y_test)
+\end{minted}
+
 % (they pass all scikit-learn's "estimators checks"). 
@@
 Weakly supervised algorithms (pair, triplet and quadruplet learners) \texttt{fit} and \texttt{predict} on a set of tuples passed as a 3-dimensional array.
-Tuples can be pairs triplets, or quadruplets depending on the algorithm.
+Tuples can be pairs, triplets, or quadruplets depending on the algorithm.
 Pair learners take as input an array-like \texttt{pairs} of shape \texttt{(n\_pairs, 2, n\_features)}, as well as an array-like \texttt{y\_pairs} of shape \texttt{(n\_pairs,)} giving labels (similar or dissimilar) for each pair.
@@
 This threshold can be set manually or automatically calibrated (at fit time or afterwards on a validation set) to optimize a given score such as accuracy or F1-score using the method \texttt{calibrate\_threshold}.
-Triplet learners work on array-like of shape \texttt{(n\_quadruplets, 3, n\_features)}, where for each triplet the second element is the one we want to be closer to the first than the third one.
-Quadruplet learners work on array-like of shape \texttt{(n\_quadruplets, 4, n\_features)}, where for each quadruplet the two first elements are the ones we want to be closer than the two last ones.
+Triplet learners work on array-like of shape \texttt{(n\_triplets, 3, n\_features)}, where for each triplet we want the first element to be closer to the second than to the third one.
+Quadruplet learners work on array-like of shape \texttt{(n\_quadruplets, 4, n\_features)}, where for each quadruplet we want the two first elements to be closer together than the two last ones.
 Both triplet and quadruplet learners can naturally \texttt{predict} whether a new triplet/quadruplet is in the right order by comparing the two pairwise distances.
-These design choices enable use of \texttt{scikit-learn}'s scoring functions out of the box, as well as the standard routines for model selection, including \texttt{GridSearchCV}.
-To illustrate, the following code snippet computes cross validation scores for ITML (with default parameters) on pairs from Labeled Faces in the Wild \citep{Huang12}.
-% Also note that for every pair/quadruplet learner, there is a \texttt{\_Supervised} version (e.g.
-for \texttt{MMC} there is \texttt{MMC\_Supervised}, that is a supervised version that samples pairs of similar samples from intra-class samples, and dissimilar samples from inter-class samples, and fits \texttt{MMC} on these pairs).
+To illustrate the weakly-supervised learning API, the following code snippet computes cross validation scores for MMC on pairs from Labeled Faces in the Wild \citep{Huang12}.
+Thanks to our unified interface, MMC can be switched for another pair learner wihtout changing the rest of the code below.
 \begin{minted}[%linenos=TRUE,
 fontsize=\footnotesize]{python}
->>> from sklearn.datasets import fetch_lfw_pairs 
->>> from sklearn.model_selection import cross_validate, train_test_split 
->>> from sklearn.decomposition import PCA 
->>> from metric_learn import ITML 
->>> pairs, y_pairs = [fetch_lfw_pairs()[key] for key in ['pairs', 'target']]
->>> pairs = PCA(n_components=25).fit_transform(pairs.reshape(4400, -1)).reshape(-1, 2, 25) 
->>> pairs, _, y_pairs, _ = train_test_split(pairs, 2*y_pairs-1)  
->>> cross_validate(ITML(), pairs, y_pairs, scoring='roc_auc', return_train_score=True)
+from sklearn.datasets import fetch_lfw_pairs 
+from sklearn.model_selection import cross_validate, train_test_split 
+from metric_learn import MMC
+
+ds = fetch_lfw_pairs()
+pairs = ds.pairs.reshape(*ds.pairs.shape[:2], -1)  # we transform 2D images into 1D vectors
+y_pairs = 2 * ds.target - 1  # we need the labels to be in {+1, -1}
+pairs, _, y_pairs, _ = train_test_split(pairs, y_pairs)
+cross_validate(MMC(diagonal=True), pairs, y_pairs, scoring='roc_auc', 
+               return_train_score=True, cv=3, n_jobs=-1, verbose=True) 
 \end{minted}
+%These design choices enable the use of \texttt{scikit-learn}'s scoring functions out of the box, as well as of the standard routines for model selection, including \texttt{GridSearchCV}.
+
 % mmc.predict(pairs)
@@
 % We would also like to offer better support to inputs in sparse format.
-We also plan to incorporate new algorithms that provide added value to the package, in particular some that learn from triplet supervision \citep{Schultz2003a}, can deal with multi-label \citep{liu15} and high-dimensional problems \citep{Liu19}, or learn other forms of metrics \citep[e.g., nonlinear ones, bilinear similarities, and multiple local metrics, see][]{Bellet15}.
+We also plan to incorporate recent algorithms that provide added value to the package, such as those that can deal with multi-label \citep{liu15} and high-dimensional problems \citep{Liu19}, or learn other forms of metrics like bilinear similarities, nonlinear and local metrics \citep[see][for a survey]{Bellet15}.
 %We could also allow metric learn to be used when having multiple labels for each sample, and allow different kind of supervisions, like semi-supervised learning for instance.

### 2020-07-03 metric-learn-jmlr.tex
Updates from Overleaf

@@
 
-% Heading arguments are {volume}{year}{pages}{submitted}{published}{author-full-names}
+\usepackage{lastpage}
+\jmlrheading{21}{2020}{1-\pageref{LastPage}}{8/19; Revised
+7/20}{7/20}{19-678}{William de Vazelhes and CJ Carey and Yuan Tang and Nathalie Vauquier and Aur\'elien Bellet}
+\ShortHeadings{metric-learn: Metric Learning Algorithms in Python}{de Vazelhes, Carey, Tang, Vauquier and Bellet}
 
-\jmlrheading{}{}{}{}{}{William de Vazelhes and CJ Carey and Yuan Tang and Nathalie Vauquier and Aur\'elien Bellet}
-
-% Short headings should be running head and authors last names
-
-\ShortHeadings{metric-learn: Metric Learning Algorithms in Python}{de Vazelhes, Carey, Tang, Vauquier and Bellet}
 \firstpageno{1}
@@
 
-\author{\name William de Vazelhes \email william.de-vazelhes@inria.fr \\
-       \addr INRIA, France
+\author{\name William de Vazelhes\thanks{Most of the work was carried out while the author was affiliated with INRIA, France.} \email wdevazelhes@gmail.com \\
+       \addr Paris Research Center, Huawei Technologies\\
+       92100 Boulogne-Billancourt, France
        \AND
@@
        \name Yuan Tang \email terrytangyuan@gmail.com \\
-       \addr Ant Financial Services Group, United States
+       \addr Ant Group, United States
        \AND
        \name Nathalie Vauquier \email nathalie.vauquier@inria.fr \\
-       \addr INRIA, France
-       \AND
+       % \addr INRIA, France
        \name Aur\'elien Bellet \email aurelien.bellet@inria.fr \\
-       \addr INRIA, France
+       \addr Magnet Team, INRIA Lille – Nord Europe\\
+       59650 Villeneuve d’Ascq, France
        }
 
-\editor{}
+\editor{Balazs Kegl}
 
@@
 \paragraph{Positioning with respect to other packages.} Many metric learning algorithms were originally implemented by their authors in \MATLAB without a common API convention.\footnote{See \url{https://www.cs.cmu.edu/~liuy/distlearn.htm} for a list of \MATLAB implementations.}  In R, the package \texttt{dml} \citep{Tang18} implements several metric learning algorithms with a unified interface but is not tightly integrated with any general-purpose machine learning library.
-In Python, \texttt{pyDML}\footnote{\url{http://github.com/jlsuarezdiaz/pydml}} contains mainly fully supervised and unsupervised algorithms, while \texttt{pytorch-metric-learning}\footnote{\url{http://github.com/KevinMusgrave/pytorch-metric-learning}} focuses on deep metric learning using the \texttt{pytorch} framework \citep{pytorch}.
+In Python, \texttt{pyDML} \citep{pyDML} contains mainly fully supervised and unsupervised algorithms, while \texttt{pytorch-metric-learning}\footnote{\url{http://github.com/KevinMusgrave/pytorch-metric-learning}} focuses on deep metric learning using the \texttt{pytorch} framework \citep{pytorch}.
 
@@
 
-The current release of \texttt{metric-learn} (v0.6.0) can be installed from the Python Package Index (PyPI) and conda-forge, for Python 3.6 or later.\footnote{Support for Python 2.7 and 3.5 was dropped in the last release.}
+The current release of \texttt{metric-learn} (v0.6.2) can be installed from the Python Package Index (PyPI) and conda-forge, for Python 3.6 or later.\footnote{Support for Python 2.7 and 3.5 was dropped in v0.6.0.}
 The source code is available on GitHub at \url{http://github.com/scikit-learn-contrib/metric-learn} and is free to use, provided under the MIT license. 

### 2020-07-06 19-678.tex
Updates from Overleaf

--- 2020-07-03 metric-learn-jmlr.tex
@@
 
-\usepackage{jmlr2e}
 \usepackage[cachedir=.,finalizecache]{minted}
@@
 \usepackage[colorinlistoftodos]{todonotes}
+\usepackage{jmlr2e}
 \newcommand{\aurelien}[1]{\todo[inline,caption={},color=orange!40]{{\it Aurelien:~}#1}}
@@
 \newcommand{\fracpartial}[2]{\frac{\partial #1}{\partial  #2}}
-
+\abbrvbib
 \usepackage{lastpage}
@@
        \name CJ Carey \email perimosocordiae@gmail.com \\
-       \addr Google LLC, United States
+       \addr Google LLC \\
+       111 8th Ave, New York, NY 10011, USA
        \AND
        \name Yuan Tang \email terrytangyuan@gmail.com \\
-       \addr Ant Group, United States
+       \addr Ant Group \\ 
+       525 Almanor Ave, Sunnyvale, CA 94085, USA
        \AND
@@
 (\subref{fig:quadruplets}) Quadruplet supervision: make leftmost pairs (here, people of similar age) closer than rightmost pairs (of bigger age difference).
-    illustrated on face image data taken from the Labeled Faces in the Wild dataset \citep{Huang12}.}\label{fig:flowers}
+    illustrated on face image data taken from the Labeled Faces in the Wild data set \citep{Huang12}.}\label{fig:flowers}
 \end{figure}
@@
 \texttt{metric-learn} currently implements algorithms that fall into the following categories.
-\emph{Supervised learners} learn from a dataset with one label per training example, aiming to bring together points from the same class while spreading points from different classes.
+\emph{Supervised learners} learn from a data set with one label per training example, aiming to bring together points from the same class while spreading points from different classes.
 For instance, data points could be face images and the class could be the identity of the person (see Figure~\ref{fig:full}). 
@@
 (\subref{fig:quadruplets}) Quadruplet supervision: make the two leftmost samples (here, people of similar age) to be closer than the two rightmost samples (of bigger age difference).
-Images are taken from the Labeled Faces in the Wild dataset \citep{Huang12}.}
+Images are taken from the Labeled Faces in the Wild data set \citep{Huang12}.}
 %   \label{fig:contour}
@@
 A distance function can also be extracted from them using \texttt{get\_metric}, that can further be used in any scikit-learn estimator like \texttt{KMeansClustering}.
-They can also return the distance between a dataset of pairs (in the 3D array format (see below), using \texttt{score\_pairs}.
+They can also return the distance between a data set of pairs (in the 3D array format (see below), using \texttt{score\_pairs}.
 
@@
 As such, they are compatible for pipelining with other estimators via \texttt{sklearn.pipeline.Pipeline}.
-To illustrate, the following code snippet trains a \texttt{Pipeline} composed of LMNN followed by a k-nearest neighbors classifier on the UCI Wine dataset, with the hyperparameters selected with a grid-search.
+To illustrate, the following code snippet trains a \texttt{Pipeline} composed of LMNN followed by a k-nearest neighbors classifier on the UCI Wine data set, with the hyperparameters selected with a grid-search.
 Any other supervised metric learner can be used in place of LMNN.
@@
 %fontsize=\footnotesize]{python}
-%>>> from sklearn.datasets import fetch_lfw_pairs 
+%>>> from sklearn.data sets import fetch_lfw_pairs 
 %>>> from sklearn.model_selection import cross_validate, train_test_split 
@@
 from metric_learn import LMNN
-
 X_train, X_test, y_train, y_test = train_test_split(*load_wine(return_X_y=True))
@@
 from metric_learn import MMC
-
 ds = fetch_lfw_pairs()
@@
 We list here some promising directions to further improve the package.
-To scale to large datasets, we would like to implement stochastic solvers (SGD and its variants), forming batches of tuples on the fly to avoid loading all data in memory at once.
+To scale to large data sets, we would like to implement stochastic solvers (SGD and its variants), forming batches of tuples on the fly to avoid loading all data in memory at once.
 % We would also like to offer better support to inputs in sparse format.
@@
 %We could also allow metric learn to be used when having multiple labels for each sample, and allow different kind of supervisions, like semi-supervised learning for instance.
-%We could also improve the ease of use by adding more examples and adding toy datasets to play with quickly and see the interest of the weaker forms of supervision.
+%We could also improve the ease of use by adding more examples and adding toy data sets to play with quickly and see the interest of the weaker forms of supervision.
 About the overall quality of the software, we intend to improve the continuous integration routines, for instance by ensuring a PEP8 syntax checker is run on every pull request, a continuous integration for the documentation.

### 2020-07-07 19-678.tex
Updates from Overleaf

@@
 \usepackage[colorinlistoftodos]{todonotes}
-\usepackage{jmlr2e}
+\usepackage[abbrvbib]{jmlr2e} % I moved this here because figure caption did not render well see the last point of http://www.jmlr.org/format/formatting-errors.html
 \newcommand{\aurelien}[1]{\todo[inline,caption={},color=orange!40]{{\it Aurelien:~}#1}}
@@
 \newcommand{\fracpartial}[2]{\frac{\partial #1}{\partial  #2}}
-\abbrvbib
+
 \usepackage{lastpage}
@@
 \begin{keywords}
-  Machine Learning, Python, Metric Learning, Scikit-learn
+  machine learning, python, metric learning, scikit-learn
 \end{keywords}
@@
 \acks
-
 We are thankful to Inria for funding 2 years of development.

### 2020-07-07 response.tex
Put back tex files

\documentclass[a4paper,12pt]{article}

\input{header-PBP.tex}
%\setlength\parindent{0pt}

%%%%%% CROSS-REFERENCING WITH THE PAPER
%\usepackage{zref-xr, zref-user}
%\zexternaldocument*{C:/Users/Emilie/Documents/Main/Recherche/EnCours/S-EmpProcSurSamp/PAPER/ROUND2/EPSS}
% use \zref{} instead of \ref{} to reference the external document(s)

%%%%% DOCUMENT INFORMATIONS

% Authors (short version in brackets, optional)
\author{W. de Vazelhes, CJ Carey, Y. Tang, N. Vauquier, A. Bellet}

% Title of the submitted paper (short version in brackets, optional)
\paper{metric-learn: Metric Learning Algorithms in Python}

% Name of the journal (short version in brackets, optional)
\journal[JMLR]{Journal of Machine Learning Research}

% Submission number
\subnb{19-678(1)}

% Date
\date{\today}

\begin{document}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Description of the revised version}

We thank the referee for his/her helpful comments. As shown by the following
point-by-point response, we have addressed all the comments. Additionally, we have updated the content of the paper to reflect the
improvements made in the last release of \texttt{metric-learn} (v0.6.0, July
2020), in particular the addition of a new algorithm (Sparse Compositional Metric Learning) along with a generic API for triplet-based metric learning algorithms.\footnote{See \url{https://github.com/scikit-learn-contrib/metric-learn/releases} for a full list of changes.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\section{Reply to the first referee}

\begin{enumerate}
    \item \textbf{Referee:} There are other metric learning libraries in various programming languages. Perharps a brief comparison with these libraries, specifying similarities and differences, may be of interest to highlight the novelty of this contribution.

    {\it \textbf{Response:} Thank you for the suggestion. We have added a
    paragraph on the positioning with respect to other packages at the end of
    Section~1.}

	\item \textbf{Referee:} Perhaps, as for the proposed code example, I would put one more general example that better captures the essence of what the software is, I see the current example a bit specific and it might be better in the user documentation rather than in the four pages description.

    {\it \textbf{Response:} In the revised version of the paper, we provide
    two separate code snippets. The first one covers supervised metric learning
    and showcases the advantages of our unified interface and tight
    integration with \texttt{scikit-learn} through the use of grid search and
    pipelining with another estimator from \texttt{scikit-learn}. The second
    snippet covers weakly supervised algorithms and is a simplified version of
    the one we had in the initial version of the paper. In both cases, we
    emphasize that the metric learning algorithm used in the snippet can
    be readily switched for another one thanks to our unified interface.}

    \item \textbf{Referee:} It would be interesting too to highlight the advantages of providing a unified interface.

    {\it \textbf{Response:} We have highlighted this better through
    the positioning with respect to other packages and our updated code
    snippets, see above points.}

    \item \textbf{Referee:} Most recent method implemented dates from 2012. Although it is also true that the library implements the best known and possibly most prominent methods, such as LMNN, NCA, MMC or ITML, which are all prior to that year, so I do not see it a drawback. Also in the future work is planned the implementation of more recent methods. In any case, the addition of some more recent method would be positive.

    {\it \textbf{Response:} In the last release of \texttt{metric-learn} 
    (July 2020), we have added a more recent algorithm (Sparse
    Compositional Metric Learning, AAAI 2014) along with a generic API for
    triplet-based metric learning which will make future implementations
    of such algorithms easier. We have updated the content of the paper to
    reflect these changes, see for instance the updated version of Figure~1
    which now illustrates triplet supervision. We will continue to incorporate
    recently proposed algorithms, as emphasized in the conclusion of the paper.}

\end{enumerate}

%\bibliographystyle{plain}
%\bibliography{EPSS}

\end{document}

### 2020-07-27 19-678.tex
Updates from Overleaf

@@
 As part of \texttt{scikit-learn-contrib}, it provides a unified interface compatible with \texttt{scikit-learn} which allows to easily perform cross-validation, model selection, and pipelining with other machine learning estimators.
-\texttt{metric-learn} is thoroughly tested and available on PyPi under the MIT licence.
+\texttt{metric-learn} is thoroughly tested and available on PyPi under the MIT license.
 %The source code is available online at \url{http://github.com/scikit-learn-contrib/metric-learn}, as well as the documentation at \url{http://metric-learn.github.io/metric-learn/}.
@@
 To illustrate the weakly-supervised learning API, the following code snippet computes cross validation scores for MMC on pairs from Labeled Faces in the Wild \citep{Huang12}.
-Thanks to our unified interface, MMC can be switched for another pair learner wihtout changing the rest of the code below.
+Thanks to our unified interface, MMC can be switched for another pair learner without changing the rest of the code below.
 \begin{minted}[%linenos=TRUE,
